---
title: "Replicating Table 1"
date: "Last compiled on `r format(Sys.time(), '%B, %Y')`"
bibliography: references.bib
output:
  html_document:
    css: tweaks.css
    toc: false
    toc_float: false
    collapsed: false
    number_sections: false
    toc_depth: 1
    code_folding: show
    code_download: yes
---


```{r, globalsettings, echo=FALSE, warning=FALSE}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```



---

This script replicates Table 1 of the manuscript (Descriptive statistics of Strava Clubs).

<br>


Clean the working environment, load in the club (RSiena) data, load some packages.

```{r, warning=FALSE}
# clean the working environment 
rm (list = ls( ))

# load the RSiena objects
load("clubdata_rsiena_freq.RData")
load("clubdata_rsiena_vol.RData")
load("clubdata.RData")

library(dplyr)
library(RSiena)
library(sna)
library(igraph)
```

<br>

Define some functions

```{r eval=T}
# density: observed relations divided by possible relations
fdensity <- function(x) {
    # x is your nomination network make sure diagonal cells are NA
    diag(x) <- NA
    # take care of RSiena structural zeros, set as missing.
    x[x == 10] <- NA
    sum(x == 1, na.rm = T)/(sum(x == 1 | x == 0, na.rm = T))
}

# calculate intragroup density
fdensityintra <- function(x, A) {
    # A is matrix indicating whether nodes in dyad have same node attributes
    diag(x) <- NA
    x[x == 10] <- NA
    diag(A) <- NA
    sum(x == 1 & A == 1, na.rm = T)/(sum((x == 1 | x == 0) & A == 1, na.rm = T))
}

# calculate intragroup density
fdensityinter <- function(x, A) {
    # A is matrix indicating whether nodes in dyad have same node attributes
    diag(x) <- NA
    x[x == 10] <- NA
    diag(A) <- NA
    sum(x == 1 & A != 1, na.rm = T)/(sum((x == 1 | x == 0) & A != 1, na.rm = T))
}

# construct dyad characteristic whether nodes are similar/homogenous
fhomomat <- function(x) {
    # x is a vector of node-covariate
    xmat <- matrix(x, nrow = length(x), ncol = length(x))
    xmatt <- t(xmat)
    xhomo <- xmat == xmatt
    return(xhomo)
}

# a function to calculate all valid dyads.
fndyads <- function(x) {
    diag(x) <- NA
    x[x == 10] <- NA
    (sum((x == 1 | x == 0), na.rm = T))
}

# a function to calculate all valid intragroupdyads.
fndyads2 <- function(x, A) {
    diag(x) <- NA
    x[x == 10] <- NA
    diag(A) <- NA
    (sum((x == 1 | x == 0) & A == 1, na.rm = T))
}


fscolnet <- function(network, ccovar) {
    # Calculate coleman on network level:
    # https://reader.elsevier.com/reader/sd/pii/S0378873314000239?token=A42F99FF6E2B750436DD2CB0DB7B1F41BDEC16052A45683C02644DAF88215A3379636B2AA197B65941D6373E9E2EE413
    
    fhomomat <- function(x) {
        xmat <- matrix(x, nrow = length(x), ncol = length(x))
        xmatt <- t(xmat)
        xhomo <- xmat == xmatt
        return(xhomo)
    }
    
    fsumintra <- function(x, A) {
        # A is matrix indicating whether nodes constituting dyad have same characteristics
        diag(x) <- NA
        x[x == 10] <- NA
        diag(A) <- NA
        sum(x == 1 & A == 1, na.rm = T)
    }
    
    # expecation w*=sum_g sum_i (ni((ng-1)/(N-1)))
    network[network == 10] <- NA
    ni <- rowSums(network, na.rm = T)
    ng <- NA
    for (i in 1:length(ccovar)) {
        ng[i] <- table(ccovar)[rownames(table(ccovar)) == ccovar[i]]
    }
    N <- length(ccovar)
    wexp <- sum(ni * ((ng - 1)/(N - 1)), na.rm = T)
    
    # wgg1 how many intragroup ties
    w <- fsumintra(network, fhomomat(ccovar))
    
    Scol_net <- ifelse(w >= wexp, (w - wexp)/(sum(ni, na.rm = T) - wexp), (w - wexp)/wexp)
    return(Scol_net)
}

fMoran.I <- function(x, weight, scaled = FALSE, na.rm = FALSE, alternative = "two.sided", rowstandardize = TRUE) {
    if (rowstandardize) {
        if (dim(weight)[1] != dim(weight)[2]) 
            stop("'weight' must be a square matrix")
        n <- length(x)
        if (dim(weight)[1] != n) 
            stop("'weight' must have as many rows as observations in 'x'")
        ei <- -1/(n - 1)
        nas <- is.na(x)
        if (any(nas)) {
            if (na.rm) {
                x <- x[!nas]
                n <- length(x)
                weight <- weight[!nas, !nas]
            } else {
                warning("'x' has missing values: maybe you wanted to set na.rm = TRUE?")
                return(list(observed = NA, expected = ei, sd = NA, p.value = NA))
            }
        }
        ROWSUM <- rowSums(weight)
        ROWSUM[ROWSUM == 0] <- 1
        weight <- weight/ROWSUM
        s <- sum(weight)
        m <- mean(x)
        y <- x - m
        cv <- sum(weight * y %o% y)
        v <- sum(y^2)
        obs <- (n/s) * (cv/v)
        if (scaled) {
            i.max <- (n/s) * (sd(rowSums(weight) * y)/sqrt(v/(n - 1)))
            obs <- obs/i.max
        }
        S1 <- 0.5 * sum((weight + t(weight))^2)
        S2 <- sum((apply(weight, 1, sum) + apply(weight, 2, sum))^2)
        s.sq <- s^2
        k <- (sum(y^4)/n)/(v/n)^2
        sdi <- sqrt((n * ((n^2 - 3 * n + 3) * S1 - n * S2 + 3 * s.sq) - k * (n * (n - 1) * S1 - 2 * n * 
            S2 + 6 * s.sq))/((n - 1) * (n - 2) * (n - 3) * s.sq) - 1/((n - 1)^2))
        alternative <- match.arg(alternative, c("two.sided", "less", "greater"))
        pv <- pnorm(obs, mean = ei, sd = sdi)
        if (alternative == "two.sided") 
            pv <- if (obs <= ei) 
                2 * pv else 2 * (1 - pv)
        if (alternative == "greater") 
            pv <- 1 - pv
        list(observed = obs, expected = ei, sd = sdi, p.value = pv)
    } else {
        if (dim(weight)[1] != dim(weight)[2]) 
            stop("'weight' must be a square matrix")
        n <- length(x)
        if (dim(weight)[1] != n) 
            stop("'weight' must have as many rows as observations in 'x'")
        ei <- -1/(n - 1)
        nas <- is.na(x)
        if (any(nas)) {
            if (na.rm) {
                x <- x[!nas]
                n <- length(x)
                weight <- weight[!nas, !nas]
            } else {
                warning("'x' has missing values: maybe you wanted to set na.rm = TRUE?")
                return(list(observed = NA, expected = ei, sd = NA, p.value = NA))
            }
        }
        # ROWSUM <- rowSums(weight) ROWSUM[ROWSUM == 0] <- 1 weight <- weight/ROWSUM
        s <- sum(weight)
        m <- mean(x)
        y <- x - m
        cv <- sum(weight * y %o% y)
        v <- sum(y^2)
        obs <- (n/s) * (cv/v)
        if (scaled) {
            i.max <- (n/s) * (sd(rowSums(weight) * y)/sqrt(v/(n - 1)))
            obs <- obs/i.max
        }
        S1 <- 0.5 * sum((weight + t(weight))^2)
        S2 <- sum((apply(weight, 1, sum) + apply(weight, 2, sum))^2)
        s.sq <- s^2
        k <- (sum(y^4)/n)/(v/n)^2
        sdi <- sqrt((n * ((n^2 - 3 * n + 3) * S1 - n * S2 + 3 * s.sq) - k * (n * (n - 1) * S1 - 2 * n * 
            S2 + 6 * s.sq))/((n - 1) * (n - 2) * (n - 3) * s.sq) - 1/((n - 1)^2))
        alternative <- match.arg(alternative, c("two.sided", "less", "greater"))
        pv <- pnorm(obs, mean = ei, sd = sdi)
        if (alternative == "two.sided") 
            pv <- if (obs <= ei) 
                2 * pv else 2 * (1 - pv)
        if (alternative == "greater") 
            pv <- 1 - pv
        list(observed = obs, expected = ei, sd = sdi, p.value = pv)
    }
    
    
}
```

<br>

For network measures, mean values over time are given with standard deviations in parentheses. For behavioral attributes, means and standard deviations at baseline (t1: December) are given. Degrees reflect kudos ties, with awarding 1 or more kudos constituting the presence of a tie.

```{r eval=T}
# Table
tab <- matrix(nrow=13, ncol=5)
rownames(tab) <- c("Density", "Average degree", "Reciprocity index", "Global transitivity index", "Jaccard index", "Gender segregation", "Moran's I: direct ties (frequency)", "Moran's I: direct ties (volume)", "Moran's I: (in)direct ties (frequency)", "Moran's I: (in)direct ties (volume)", "Weekly running frequency at t1", "Weekly running hours at t1", "Male sex (%)" )
colnames(tab) <- c("Club 1","Club 2", "Club 3", "Club 4", "Club 5")

for (i in 1:5) { # for each club
  # densities over time
  densities <- as.numeric(strsplit(substring(capture.output(clubdata_rsiena_freq[[i]])[11], 20), " ")[[1]])
  # mean (standard deviation)
  tab[1, i] <- paste(as.character(round(mean(densities), 3)), " (", as.character(round(sd(densities), 3)), ")", sep="")
}

# Average degrees
for (i in 1:5) { # for each club
  # calculate average out-degree over time
  d <- vector ()
  for (t in 1:12) {
    d[[t]] <- mean(igraph::degree(igraph::graph_from_adjacency_matrix(clubdata_rsiena_freq[[i]]$depvars$kudonet[,,t]), mode = "out"))

  }
  # mean (standard deviation)
  tab[2, i] <- paste(as.character(round(mean(d), 3)), " (", as.character(round(sd(d), 3)), ")", sep="")
}

# Reciprocity index
for (i in 1:5) { # for each club
  # calculate reciprocity index over time
  ri <- vector ()

  for (t in 1:12) {
    mut <- igraph::dyad.census(igraph::graph_from_adjacency_matrix(clubdata_rsiena_freq[[i]]$depvars$kudonet[,,t]))[[1]]
    nonmut <- igraph::dyad.census(igraph::graph_from_adjacency_matrix(clubdata_rsiena_freq[[i]]$depvars$kudonet[,,t]))[[2]]
    ri[[t]] <- mut/(mut+nonmut)
    
  }
  tab[3, i] <- paste(as.character(round(mean(ri), 3)), " (", as.character(round(sd(ri), 3)), ")", sep="")
}

# Transitivity index
for (i in 1:5) { # for each club
  # calculate transitivity index over time
  ti <- vector ()

  for (t in 1:12) {
    ti[[t]] <- igraph::transitivity(igraph::graph_from_adjacency_matrix(clubdata_rsiena_freq[[i]]$depvars$kudonet[,,t]))
    
  }
  tab[4, i] <- paste(as.character(round(mean(ti), 3)), " (", as.character(round(sd(ti), 3)), ")", sep="")
}

# Jaccard index

# Gender segregation 
for (i in 1:5) { # for each club
  # calculate coleman's homophily index over time
  col <- vector ()
  
  for (t in 1:12) {
    col[[t]] <- fscolnet(clubdata_rsiena_freq[[i]]$depvars$kudonet[,,t], clubdata_rsiena_freq[[i]]$cCovars$gender)
  }
  tab[6, i] <- paste(as.character(round(mean(col), 3)), " (", as.character(round(sd(col), 3)), ")", sep="")
}

# Network autocorrelation
# frequency
for (i in 1:5) { # for each club
  # we calculate moran's i
  # both with direct ties and indirect ties as well
  # for frequency and volume
  mor_freq_dir <- vector ()
  mor_freq_dist <- vector ()
  mor_vol_dir <- vector ()
  mor_vol_dist <- vector ()
  
  for (t in 1:12) {         
    freq <- clubdata[[i]]$freq_run[,,t]   # attributes at time t
    vol <- clubdata[[i]]$time_run[,,t]   
    knet <- network::as.network(clubdata_rsiena_freq[[i]]$depvars$kudonet[,,t]) # net a time t
    geodistances <- sna::geodist(knet, count.paths=T) # geodesic distance
    geodistances <- geodistances$gdist 
    diag(geodistances) <- Inf
    weights1 <- geodistances == 1               # only direct ties
    weights2 <- exp(-geodistances)              # also indirect
    
    mor_freq_dir[[t]] <- fMoran.I(freq, scaled = FALSE, # calculate moran on direct
                       weight = weights1, 
                       na.rm = TRUE, 
                       rowstandardize = FALSE)[[1]] 
    mor_freq_dist[[t]] <- fMoran.I(freq, scaled = FALSE, # and indirect
                       weight = weights2, 
                       na.rm = TRUE, 
                       rowstandardize = FALSE)[[1]]
    mor_vol_dir[[t]] <- fMoran.I(vol, scaled = FALSE, # calculate moran on direct
                       weight = weights1, 
                       na.rm = TRUE, 
                       rowstandardize = FALSE)[[1]] 
    mor_vol_dist[[t]] <- fMoran.I(vol, scaled = FALSE, # and indirect
                       weight = weights2, 
                       na.rm = TRUE, 
                       rowstandardize = FALSE)[[1]]
  }
  tab[7, i] <- paste(as.character(round(mean(mor_freq_dir), 3)), " (", as.character(round(sd(mor_freq_dir), 3)), ")", sep="")
  tab[8, i] <- paste(as.character(round(mean(mor_vol_dir), 3)), " (", as.character(round(sd(mor_vol_dir), 3)), ")", sep="")
  tab[9, i] <- paste(as.character(round(mean(mor_freq_dist), 3)), " (", as.character(round(sd(mor_freq_dist), 3)), ")", sep="")
  tab[10, i] <- paste(as.character(round(mean(mor_vol_dist), 3)), " (", as.character(round(sd(mor_vol_dist), 3)), ")", sep="")
}

# behavior at baseline (w1)
for (i in 1:5) { # for each club
  # running attributes at w1
  tab[11, i] <- paste(as.character(round(mean(clubdata[[i]]$freq_run[,,1]), 3)), " (", as.character(round(sd(clubdata[[i]]$freq_run[,,1]), 3)), ")", sep="")
  tab[12, i] <- paste(as.character(round(mean(clubdata[[i]]$time_run[,,1]), 3)), " (", as.character(round(sd(clubdata[[i]]$time_run[,,1]), 3)), ")", sep="")
}
# male sex (%)
for (i in 1:5) { # for each club
  # percentage male
  tab[13, i] <- round((length(clubdata[[i]][clubdata[[i]]$male==1])/clubdata[[i]]$netsize)*100, 3)
}

``` 
  

<b>



```{r table}

knitr::kable(tab, digits=3, "html", caption="Table 1. Descriptive statistics of Strava clubs") %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) 

```

