---
title: "RSiena: Exploratory analyses "
date: "Last compiled on `r format(Sys.time(), '%B, %Y')`"
bibliography: references.bib
output:
  html_document:
    css: tweaks.css
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    code_folding: show
    code_download: yes
---

```{r, globalsettings, echo=FALSE, warning=FALSE}
library(knitr)
library(RSiena)
knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```



---

Let's estimate the Stochastic Actor-Oriented Model (SAOM) implemented in R as the Simulation Investigation for Empirical Network Analysis (R-SIENA), developed by @snijders2010.

----

<br>

# Preparation

Clean the working environment.

```{r, attr.output='style="max-height: 200px;"'}
# clean the working environment 
rm (list = ls( ))
```

<br> 

On this page we will:

1. Get our data and create an R-SIENA object
2. Inspect our data
3. Define our effects
4. Define our algorithm
5. And estimate the SAOM

Below, we will follow these steps for club 1 (N=30). We will explore the SAOM for 3 clubs in order to find the best model (see [Results](#result)). We will subsequently test if this model also fits the data of the remaining two clubs (and afterwards with the ego-centered networks).


----

<br>

# Step 1: Data

We start off by reading in the data, specifying variable roles for the RSiena object, and creating an RSiena object. We use the clubdata.RData object.

- For now we take as our (explanatory) network variable the Kudo-network in which awarding/receiving *at least* 1 Kudo constitutes an *i,j* tie. Later on, we may want to use the ordered variable.

- Our (dependent) behavioral variable is running frequency (in times per week; ranging from 0 to >7 times per week), and running volume (in minutes per month; with 8 equally populated categories).

- We included activity in other sports (e.g., cycling and swimming) as a time-varying covariate.

- And we also included gender (men vs. women and others), and winter as constant and varying covariates, respectively, and friendship as a time-constant dyadic covariate. 


```{r}
load("clubdata.Rdata") # load (raw) club list
club <- clubdata[[1]] # grab club 

# specify variable roles for RSiena object
kudonet <- sienaDependent(club$kudo)

freq_run <- sienaDependent(club$freq_run, type= "behavior")

time_run <- sienaDependent(club$time_run, type = "behavior")

freq_other <- varCovar(club$freq_other[,,])
time_other <- varCovar(club$time_other[,,])

gender <- NA 
gender <- ifelse(club$male == 1, 1, gender)
gender <- ifelse(club$female == 1, 2, gender)
gender <- ifelse(club$other == 1, 2, gender)
gender <- coCovar(gender)

winter <- varCovar(club$winter)

friendship <- coDyadCovar(club$friendship)

#create rsiena data object
mydata <- sienaDataCreate(kudonet, freq_run, time_run, freq_other, time_other, gender, winter, friendship)
```

----

<br>

# Step 2: Inspect data
```{r eval=T}
print01Report(mydata)
```

A text file is printed in the working directory.

![](files/Siena.txt){#id .class width=100% height=200px}

----

<br>

# Step 3: Define effects
We are going to define our 'myeff' object containing the model parameters. A list of all available effects for the given object can be displayed in browser by requesting effectsDocumentation(myeff). See @rsienamanual for a substantial and mathematical description of all effects.

We include:

1. [structural network effects](#str)
2. [network selection effects](#sel)
3. [network influence effects](#inf)
4. and [covariate effects](#co) on network and behavior 


```{r echo=T, results='hide'}
myeff <- getEffects(mydata)
#effectsDocumentation(myeff)
```

<br>

## Structural network effects {#str}
First, we are going to include structural network effects, guided by recommendations of @snijderspres: transitivity (GWESP), indegree popularity (sqrt.), outdegree activity (sqrt.), and the transitive reciprocated triplets effect. For the last effect, see @block2015.

```{r echo=T, results='hide'}
myeff1 <- includeEffects(myeff, gwespFF, 
                         inPopSqrt, outActSqrt, name = "kudonet") 
myeff1 <- includeInteraction(myeff1, recip, gwespFF, parameter = 69) 

```

<br>

## Selection effects {#sel}
Second, we include selection effects with respect to behavior. We estimate the effect of actors' (ego and alter) behavior, and the similarity between the two (homophily) on tie formation. 

We start with effects of frequency on tie formation. Later, we will add volume.

For now we only include the monadic effects (i.e. of frequency), to control for *structural* effects (i.e., being more active on Strava may increase one's opportunities to award and receive Kudos). After estimating this structural part, we can add the similarity effect, which reflects homophilous *preferences*, over and above the opportunity structures. 


```{r echo=T, results='hide'}
myeff2 <- includeEffects(myeff1, egoX, altX, name="kudonet", interaction1 = "freq_run")

myeff2 <- includeEffects(myeff2, egoX, altX, name="kudonet", interaction1 = "time_run")

myeff2a <- includeEffects(myeff2, simX, name="kudonet", interaction1 = "freq_run")

myeff2a <- includeEffects(myeff2a, simX, name="kudonet", interaction1 = "time_run")
```

<br>

## Influence effects {#inf}
We included two influence effects to test our hypotheses on the influence of social support and social comparison processes respectively:

1. The reciprocated degree effect on behavior.
This tests the social support explanation: the more reciprocated Kudo-ties (i.e. "strong tie" friendships), the greater the probability of increasing the behavior. 

2. The average reciprocated alter effect.
This test the social comparison explanation: the tendency to assimilate to the average behavior of reciprocated alters (i.e. "strong tie" friends).

**Note** that in the average reciprocated alter effect the expected positive effect of upward comparison and the negative effect of downward comparison are assumed to even each other out. In our case this assumption is rather poor, as upward and downward comparison processes may have independent effects on ego’s tendency to increase their behavior. Therefore, we may want to find a more suitable effect. 

We start with myeff3: to test if reciprocated degrees have an influence. Afterwards, we will test myeff4: to investigate through what mechanism (support vs. comparison).

```{r echo=T, results='hide'}
myeff3 <- includeEffects(myeff2, recipDeg, name = "freq_run", interaction1 = "kudonet") 
myeff3 <- includeEffects(myeff3, recipDeg, name = "time_run", interaction1 = "kudonet")

myeff4 <- includeEffects(myeff3, avRecAlt, name = "freq_run", interaction1 = "kudonet") 
myeff4 <- includeEffects(myeff4, avRecAlt, name = "time_run", interaction1 = "kudonet") 
```

<br>

An extra theoretical thought: we may want to condition the reciprocated degree effects on current behavior, to see if they hold to a greater/lesser extent for 'novice' athletes.

```{r echo=F, eval=F, results='hide'}
myeff5 <- includeEffects(myeff4, quad, recipDeg, name = "freq_run", interaction1 = c("", "kudonet"))
myeff5 <- includeEffects(myeff4, quad, avRecAlt, name = "freq_run", interaction1 = c("", "kudonet"))
```

<br>

## Covariate effects {#co}

And last we add effects on tie and behavior changes of other variables:

- the interdependence between frequency and volume
- the interdependence between running and other sports
- winter on behavior
- gender on tie selection and behavior

```{r echo=T, results='hide'}
myeff3 <- includeEffects(myeff3, effFrom, name = "time_run", interaction1 = "freq_run")
myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "time_run")

myeff3 <- includeEffects(myeff3, effFrom, name = "time_run", interaction1 = "time_other")
myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "freq_other")

myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "winter")
myeff3 <- includeEffects(myeff3, effFrom, name = "time_run", interaction1 = "winter")

myeff3 <- includeEffects( myeff3, egoX, altX, sameX, name="kudonet", interaction1 = "gender" )
myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "gender")
myeff3 <- includeEffects(myeff3, effFrom, name = "time_run", interaction1 = "gender")
```


<br>

Now check which effects are included in the myeff object.

```{r class.source = 'fold-hide' }
options(width = 100) # ignore (this is for the html formatting)
print(myeff3)
```

## {-}

Seems allright!

----

<br>

# Step 4: Define algorithm

```{r }
myalgorithm <- sienaAlgorithmCreate(projname = "test")
```

----

<br>

# Step 5: Estimate the model

Let's estimate the SAOM.

We will save the results in a html-table using the siena.table()-function, in the 'files'-folder in our working directory. We will run the model as many times as necessary, until we get an acceptable convergence ratio (< .25).

<!--- 

I like the code below. Please save the 'final' ans objects somewhere. 
I guess a could way for replication in the end is to have all final Ans objects somewhere. and then with a specific seed rerun all models? or dont we really care about this? 

---> 


```{r eval= F }

try <- 1

ansM1 <- siena07(myalgorithm, data = mydata, effects = myeff3)
siena.table(ansM1, type="html", tstat=T, d=2, sig=T, file = paste("files", "/", "last.html", sep = ""))

# the following script lets the model re-run until we get a good convergence ratio
while (TRUE){
  
  if(ansM1$tconv.max > 0.25){
    try <- try + 1
    print(paste("Try:", try, sep=" "))
    ansM1 <- siena07( myalgorithm, data = mydata, effects = myeff3, prevAns= ansM1)
    siena.table(ansM1, type="html", tstat=T, d=2, sig=T, file = paste("files", "/", "ansM1_", try, ".html", sep=""))
    
  }else{
    siena.table(ansM1, type="html", tstat=T, d=2, sig=T, file = paste("files", "/", "final", ".html", sep=""))
    print(paste("Reached convergence ratio of ", ansM1$tconv.max, sep = ""))
    break
  }
}
```

----

<br>


# Step 6: Results {#result}

Here we present the results for club 1 (N=30). We did not get a convergence ratio <.25, but close (.29), which we accept for our exploratory SAOM. 

<!--- 

Estimate the same SAOM for club 2 and 5;
Present the results
Reflect on results
Modify SAOM; extra effects where necessary.

---> 

```{r, echo=F}
htmltools::includeHTML("files/ansM1_4.html")
```

----

## References
