---
title: "SAOM: Model selection"
date: "Last compiled on `r format(Sys.time(), '%B, %Y')`"
bibliography: references.bib
output:
  html_document:
    css: tweaks.css
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    code_folding: show
    code_download: yes
---
  
```{r, globalsettings, echo=FALSE, warning=FALSE}
library(knitr)
library(RSiena)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```



---
  
Let's estimate the Stochastic Actor-Oriented Model (SAOM) implemented in R as the Simulation Investigation for Empirical Network Analysis (R-SIENA), developed by @snijders2010.

----

<br>


# Preparation

Clean the working environment.

```{r, attr.output='style="max-height: 200px;"'}
# clean the working environment 
rm (list = ls( ))
```

<br> 

We will:

1. Read in our R-SIENA object list
2. Inspect our data
3. Define our effects
4. Define our algorithm
5. And estimate the SAOM

Below, we will follow these steps for club 1 (N=30). We will do the same procedure for the other clubs (see <a href="https://robfranken.github.io/Strava/meta.html" target="_blank"><b>here</b></a>). 

<br>

----


# Step 1: Data

We read in the R-SIENA objects list (*clubdata_rsiena.RData*) and we grab club 1 (N=30). For now we take as our network variable the kudos-network in which awarding/receiving *at least* 1 Kudo constitutes an *i,j* tie. 

- Our (dependent) behavioral variable is running frequency (in times per week; ranging from 0 to 7 times per week), and running duration (in half hours per week, capped at 7).

- We included activity in other sports (e.g., cycling and swimming) as a time-varying covariate.

- And we also included gender (men vs. women and others), and winter as constant and varying covariates, respectively.


```{r}
load("clubdata_rsiena.Rdata") # load rsiena object list
mydata <- clubdata_rsiena[[1]] # grab club 1
```

<br>

----


# Step 2: Inspect data

We inspect the R-SIENA object

```{r eval=F}
print01Report(mydata, modelname="files/test")
```

A text file is printed in the working directory.

![](files/test.txt){#id .class width=100% height=200px}

----

<br>

# Step 3: Define effects
We are going to define our 'myeff' object containing the model parameters. A list of all available effects for the given object can be displayed in browser by requesting "effectsDocumentation(myeff)". See @rsienamanual for a substantial and mathematical description of all effects.

We build increasingly complex models.

We include:

1. [structural network effects](#str)
2. [network selection effects](#sel)
3. [covariate effects](#co) on network and behavior dynamics
4. [network influence effects](#inf)

We fix these effects to 0 and test them with the score-type test @Schweinberger2012 (we test the hypothesis that the parameter estimates are not 0, other than the model assumes).


```{r echo=T, results='hide'}
myeff <- getEffects(mydata)
#effectsDocumentation(myeff)
```


<br>

## Structural network effects {#str}
First, we are going to include structural network effects, guided by recommendations of @snijderspres: outdegree, reciprocity, and transitivity (GWESP).

We also add degree-related effects: indegree-popularity and outdegree-activity (square-root versions).

We tested the out-Isolate effect (leading to not awarding kudos) and this effect was not different from 0. 

```{r echo=T, results='hide'}
myeff1 <- includeEffects(myeff, gwespFF, name = "kudonet") 
myeff1 <- includeEffects(myeff1, outActSqrt, inPopSqrt, name = "kudonet") 
myeff1 <- setEffect( myeff1, outIso, name = "kudonet", fix = TRUE, test = FALSE, initialValue = 0)
```


<br>

## Selection effects {#sel}
Second, we include a selection effect with respect to behavior:

- for our 
We use the higher-effect to control for aspirational tie-preferences (indicated by a negative parameter estimate).

```{r echo=T, results='hide'}
myeff2 <- setEffect(myeff1, higher, name = "kudonet", interaction1 = "freq_run")
```


<br>
  
## Covariate effects {#co}
  
We include effects on tie changes of gender (monadic and dyadic).

```{r echo=T, results='hide'}
myeff2 <- includeEffects(myeff2, egoX, altX, sameX, name="kudonet", interaction1 = "gender" )
```

  
<br> 
  
We have selected a rather simple model to simulate kudos tie-formation dynamics. Let's estimate the model and assess the model's GOF to additional effects that were not directly modeled: the in- and outdegree distribution and the geodesic distance distribution. We use 'returnDeps=TRUE' for keeping the simulated data (networks and behavior), for subsequent GOF assesment. We save the GOF-diagnostics in a list.

```{r eval=F, results='hide'}
myalgorithm <- sienaAlgorithmCreate(projname = "test") # first, set the SAOM algorithm 

ansM1 <- siena07(myalgorithm, data = mydata, effects = myeff2, # estimate the SAOM
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE)

# calculate GOF diagnostics
gofi <- sienaGOF(ansM1, 
                 IndegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

gofo <- sienaGOF(ansM1, 
                 OutdegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

GeodesicDistribution <- function (i, data, sims, period, groupName,
   varName, levls=c(1:5, Inf), cumulative=TRUE, ...) {
     x <- networkExtraction(i, data, sims, period, groupName, varName)
     require(sna)
     a <- sna::geodist(symmetrize(x))$gdist
     if (cumulative)
     {
       gdi <- sapply(levls, function(i){ sum(a<=i) })
     }
     else
     {
       gdi <- sapply(levls, function(i){ sum(a==i) })
     }
     names(gdi) <- as.character(levls)
     gdi
}

gofgeo <- sienaGOF(ansM1, 
                 GeodesicDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

goflist <- list(gofi, gofo, gofgeo)
save(goflist, file= paste("files", "/", "test club 1", "/", "gof1.RData", sep=""))
```

### {.tabset .tabset-fade}

#### Indegree distribution
```{r class.source = 'fold-hide'}
load("files/test club 1/gof1.RData")
plot(goflist[[1]])
```

#### Outdegree distribution
```{r class.source = 'fold-hide'}
plot(goflist[[2]])
```

#### Geodesic distance distribution
```{r class.source = 'fold-hide'}
plot(goflist[[3]])
```


### {-}

GOF is acceptable!

<br>

For subsequent meta-analysis, we need to ensure that the model specification for all our club-networks is identical. Some effects were rather important in other clubs. We fix these to 0:

```{r eval=F, results='hide'}
myeff2 <- setEffect( myeff2, outPopSqrt, name = "kudonet", fix = TRUE, test = FALSE, initialValue = 0)
myeff2 <- setEffect( myeff2, reciAct, name = "kudonet", fix = TRUE, test = FALSE, initialValue = 0)
myeff2 <- includeInteraction(myeff2, recip, gwespFF, parameter = 69, name = "kudonet")
eff1 <- myeff2[myeff2$include,]$effect1[24]
eff2 <- myeff2[myeff2$include,]$effect2[24]
myeff2 <- setEffect(myeff2, unspInt, fix=TRUE, test=FALSE, effect1=eff1, effect2=eff2)
```

<!--- 

@RF: for some reasons this gives an error when rendering the site... so set on eval=F

---> 

<br>

----

We have modeled the dynamics of kudos tie formation. Now let's model dynamics in running behaviors.

## Covariate effects
We start with effects on behavior changes of other variables.

- the interdependence between frequency and volume
- the interdependence between running and other sports
- gender on behavior
- winter on behavior (we exclude this effect, as it was not significantly different from 0)
  
```{r echo=T, results='hide'}
myeff3 <- includeEffects(myeff2, effFrom, name = "time_run", interaction1 = "freq_run")
myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "time_run")

myeff3 <- includeEffects(myeff3, effFrom, name = "time_run", interaction1 = "time_other")
myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "freq_other")

myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "gender")
myeff3 <- includeEffects(myeff3, effFrom, name = "time_run", interaction1 = "gender")

```
  
<br>

----
  
## Influence effects {#inf}
Last, we include two influence effects to test our hypotheses on the influence of social support and social comparison processes respectively:
  
1. The indegree effect on behavior.
This tests the social support explanation: the higher the indegree (our operationalization of received support), the greater the probability of increasing the behavior. 

2. The average attraction towards higher alters effect
This tests social comparison as assimilation.

We test these effects sequentially:
so first, we estimate a model with the indegree effect on behavior added to the baseline model;
second, we add the average attraction to higher alters effect. 

We also fixed-and-tested the effect of outdegree on behavior, to rule out possible confounding of the outdegree effect. Score-type test indicated that outdegree-effects on behavior were 0. 


```{r echo=T, results='hide'}
myeff3 <- includeEffects(myeff3, indeg, name = "freq_run", interaction1 = "kudonet") 
myeff3 <- includeEffects(myeff3, indeg, name = "time_run", interaction1 = "kudonet")

myeff4 <- includeEffects(myeff3, avAttHigher, name = "freq_run", interaction1 = "kudonet") 
myeff4 <- includeEffects(myeff4, avAttHigher, name = "time_run", interaction1 = "kudonet")
myeff4 <- setEffect(myeff4, outdeg, name = "freq_run", interaction1 = "kudonet", fix = TRUE, test = FALSE, initialValue = 0) 
myeff4 <- setEffect(myeff4, outdeg, name = "time_run", interaction1 = "kudonet", fix = TRUE, test = FALSE, initialValue = 0)
```

  
<br>
  
Now check which effects are included in the myeff objects.

```{r class.source = 'fold-hide' }
options(width = 100) # ignore (this is for the html formatting)
print(myeff3)
print(myeff4)
```

<br>

Seems allright!
  
----
  
# Step 4: Estimate the model
  
Let's estimate the model with the objective function for running activity comprising the shape effects and the indegree effect on behavior. We use the prevAns-option to use previous model-results as starting values for the estimation.

We will save the results in a html-table using the siena.table()-function, in the 'files'-folder in our working directory. We will run the model as many times as necessary, until we reach a convergence ratio of < .25. We use 'returnDeps=TRUE' for keeping the simulated data (networks and behavior).


```{r eval= F }
try <- 1
ansM2 <- siena07(myalgorithm, data = mydata, effects = myeff3, returnDeps=TRUE, prevAns=ansM1)

# the following script lets the model re-run until we get a good convergence ratio
while (TRUE){
  
  if(ansM2$tconv.max >= .25){
    try <- try + 1
    print(paste("Model did not converge: ", ansM2$tconv.max, sep = ""))
    print(paste("Try:", try, sep=" "))
    ansM2 <- siena07( myalgorithm, data = mydata, effects = myeff3, prevAns= ansM2, returnDeps=TRUE)
    
  }else{
    siena.table(ansM2, type="html", tstat=T, d=3, sig=T, file = paste("files", "/", "test club 1", "/", "Model 2", ".html", sep=""))
    print(paste("Reached convergence ratio of ", ansM2$tconv.max, sep = ""))
    break
  }
}
```

<br>
 
This took some time. We also save the sienaFit-object to our "Principal club"-folder.

```{r eval= F }
save(ansM2, file= paste("files", "/", "test club 1", "/", "model2.RData", sep=""))
```


<br>

Now follow the same procedure for the model comprising also the assimilation- (and outdegree-) effect. Again, we will run the model as many times as necessary, until we reach convergence. Again, we use prevAns.

```{r eval= F }
try <- 1
ansM3 <- siena07(myalgorithm, data = mydata, effects = myeff4, returnDeps=TRUE, prevAns=ansM2)

# the following script lets the model re-run until we get a good convergence ratio
while (TRUE){
  
  if(ansM3$tconv.max >= .25){
    try <- try + 1
    print(paste("Model did not converge: ", ansM3$tconv.max, sep = ""))
    print(paste("Try:", try, sep=" "))
    ansM3 <- siena07( myalgorithm, data = mydata, effects = myeff4, prevAns= ansM3, returnDeps=TRUE)
    
  }else{
    siena.table(ansM3, type="html", tstat=T, d=3, sig=T, file = paste("files", "/", "test club 1", "/", "Model 3", ".html", sep=""))
    print(paste("Reached convergence ratio of ", ansM3$tconv.max, sep = ""))
    break
  }
}

```

<br>

We save the sienaFit-object.

```{r eval= F }
save(ansM3, file= paste("files", "/", "test club 1", "/", "model3.RData", sep=""))

```

<br> 

Because we are now modeling the evolution of both the network and the attribute (running activity), we will add an additional indicator to evaluate GOF; namely, does the model capture the distribution of actors’ attribute levels over time?

```{r eval=F, results='hide'}
gofbeh <- sienaGOF(ansM3,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "freq_run")

gofbeh2 <- sienaGOF(ansM3,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "time_run")

goflist2 <- list(gofbeh, gofbeh2)
save(goflist2, file= paste("files", "/", "test club 1", "/", "gof2.RData", sep=""))
```

### {.tabset .tabset-fade}

#### Running frequency distribution
```{r class.source = 'fold-hide'}
load("files/test club 1/gof2.RData")
plot(goflist2[[1]])
```

#### Running duration distribution
```{r class.source = 'fold-hide'}
plot(goflist2[[2]])
```

### {-}



GOF is adequate for the distribution of running frequency values, but not for the running duration values...

This may be due to the operationalization of this variable...


----

<br>

# Step 5: Results {#result}

We present the SAOM results for club 1.

The first table (Model 1) presents the parameter estimates for (1) network structure; (2) network dynamics (selection); (3) behavior dynamics comprising shape effects, *indegree effect* on behavior, and covariate effects on behavior change. The second table (Model 2) adds the assimilation- and outdegree-effects on behavior.

## {.tabset .tabset-fade}

### Model 1
```{r, echo=F}
htmltools::includeHTML("files/test club 1/Model 2.html")
```

### Model 2
```{r, echo=F}
htmltools::includeHTML("files/test club 1/Model 3.html")
```

## {-}

----

We do not find a significant kudos indegree effect. We do find a significant upward assimilation effect for running frequency. 

<br>

----

## Next up

We will check whether this model configuration also converges for the [other clubs](https://robfranken.github.io/Strava/meta.html). To summarize the results over our clubs, we will perform a [meta-analysis](https://robfranken.github.io/Strava/meta.html#Meta-analysis) using a Fisher-type combination of one-tailed p-values.


----


### References
