---
title: "SAOM: Model selection"
date: "Last compiled on `r format(Sys.time(), '%B, %Y')`"
bibliography: references.bib
output:
  html_document:
    css: tweaks.css
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    code_folding: show
    code_download: yes
---
  
```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)
library(RSiena)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```



---
  
Let's estimate the Stochastic Actor-Oriented Model (SAOM) implemented in R as the Simulation Investigation for Empirical Network Analysis (R-SIENA), developed by @snijders2010.

----

<br>


# Preparation

Clean the working environment.

```{r, attr.output='style="max-height: 200px;"'}
# clean the working environment 
rm (list = ls( ))
```

<br> 

We will:

1. Read in our R-SIENA object list
2. Inspect our data
3. Define our effects
4. Define our algorithm
5. And estimate the SAOM

Below, we will follow these steps for club 1 (N=30). Here, we focus on running frequency. We did the same procedure for the other clubs.

<br>

----


# Step 1: Data

We read in the R-SIENA objects list (*clubdata_rsiena_freq.RData*) and we grab club 1 (N=30). We take as our network variable the kudos-network in which awarding/receiving *at least* 1 kudos constitutes an *i,j* tie. 

- Our (dependent) behavioral variable is running frequency (in times per week; ranging from 0 to 7+ times per week).

- We included activity (frequency) in other sports (e.g., cycling and swimming) as a time-varying covariate.

- And we also included gender (men vs. women and others) as constant covariate.


```{r}
load("clubdata_rsiena_freq.Rdata") # load rsiena object list
mydata <- clubdata_rsiena_freq[[1]] # grab club 1
```

<br>

----


# Step 2: Inspect data

We inspect the R-SIENA object

```{r eval=F}
print01Report(mydata, modelname="files/test")
```

A text file is printed in the working directory.

![](files/test.txt){#id .class width=100% height=200px}

----

<br>

# Step 3: Define effects
We are going to define our 'myeff' object containing the model parameters. A list of all available effects for the given object can be displayed in browser by requesting "effectsDocumentation(myeff)". See @rsienamanual for a substantial and mathematical description of all effects.

We build increasingly complex models.

We include:

1. [structural network effects](#str)
2. [network selection effects](#sel)
3. [covariate effects](#co) on network and behavior dynamics
4. [network influence effects](#inf)

We fix these effects to 0 and test them with the score-type test @Schweinberger2012 (we test the hypothesis that the parameter estimates are not 0, other than the model assumes).


```{r echo=T, results='hide'}
myeff <- getEffects(mydata)
#effectsDocumentation(myeff)
```


<br>

## Structural network effects {#str}
First, we are going to include structural network effects, guided by recommendations of @snijderspres: outdegree, reciprocity, and transitivity (GWESP).

We also add degree-related effects: indegree-popularity and outdegree-activity (square-root versions).

We tested the out-Isolate effect (leading to not awarding kudos) and this effect was not different from 0. 

```{r echo=T, results='hide'}
myeff1 <- includeEffects(myeff, gwespFF, name = "kudonet") 
myeff1 <- includeEffects(myeff1, outActSqrt, inPopSqrt, name = "kudonet") 
myeff1 <- setEffect( myeff1, outIso, name = "kudonet", fix = TRUE, test = FALSE, initialValue = 0)
```


<br>

## Selection effects {#sel}
Second, we include selection effects with respect to behavior: egoX, altX and simX.

In addition, we use the higher-effect to control for aspirational tie-preferences (indicated by a negative parameter estimate).

```{r echo=T, results='hide'}
myeff2 <- includeEffects(myeff1, egoX, altX, simX, higher, name = "kudonet", interaction1 = "freq_run")
```


<br>
  
## Covariate effects {#co}
  
We include effects on tie changes of gender (monadic and dyadic).

```{r echo=T, results='hide'}
myeff2 <- includeEffects(myeff2, egoX, altX, sameX, name="kudonet", interaction1 = "gender" )
```

  
<br> 
  
We have selected a rather simple model to simulate kudos tie-formation dynamics. Let's estimate the model and assess the model's GOF to additional effects that were not directly modeled: the in- and outdegree distribution and the geodesic distance distribution. We use 'returnDeps=TRUE' for keeping the simulated data (networks and behavior), for subsequent GOF assesment. We save the GOF-diagnostics in a list.

```{r eval=F, results='hide'}
myalgorithm <- sienaAlgorithmCreate(projname = "test", nsub=5, n3=5000 )
 # first, set the SAOM algorithm 

ansM1 <- siena07(myalgorithm, data = mydata, effects = myeff2, # estimate the SAOM
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE)

# calculate GOF diagnostics
gofi <- sienaGOF(ansM1, 
                 IndegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

gofo <- sienaGOF(ansM1, 
                 OutdegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

GeodesicDistribution <- function (i, data, sims, period, groupName,
   varName, levls=c(1:5, Inf), cumulative=TRUE, ...) {
     x <- networkExtraction(i, data, sims, period, groupName, varName)
     require(sna)
     a <- sna::geodist(symmetrize(x))$gdist
     if (cumulative)
     {
       gdi <- sapply(levls, function(i){ sum(a<=i) })
     }
     else
     {
       gdi <- sapply(levls, function(i){ sum(a==i) })
     }
     names(gdi) <- as.character(levls)
     gdi
}

gofgeo <- sienaGOF(ansM1, 
                 GeodesicDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

goflist <- list(gofi, gofo, gofgeo)
save(goflist, file= paste("files", "/", "test club 1", "/", "gof.RData", sep=""))
```

### {.tabset .tabset-fade}

#### Indegree distribution
```{r class.source = 'fold-hide'}
load("files/test club 1/gof.RData")
plot(goflist[[1]])
```

#### Outdegree distribution
```{r class.source = 'fold-hide'}
plot(goflist[[2]])
```

#### Geodesic distance distribution
```{r class.source = 'fold-hide'}
plot(goflist[[3]])
```


### {-}

GOF is acceptable!

<br>

For subsequent meta-analysis, we need to ensure that the model specification for all our club-networks is identical. Some effects were rather important in other clubs. We fix these to 0:

```{r eval=F, results='hide'}
myeff2 <- setEffect( myeff2, outPopSqrt, name = "kudonet", fix = TRUE, test = FALSE, initialValue = 0)
myeff2 <- setEffect( myeff2, reciAct, name = "kudonet", fix = TRUE, test = FALSE, initialValue = 0)
myeff2 <- includeInteraction(myeff2, recip, gwespFF, parameter = 69, name = "kudonet")
eff1 <- myeff2[myeff2$include,]$effect1[24]
eff2 <- myeff2[myeff2$include,]$effect2[24]
myeff2 <- setEffect(myeff2, unspInt, fix=TRUE, test=FALSE, effect1=eff1, effect2=eff2)
```

<!--- 

@RF: for some reasons this gives an error when rendering the site... so set on eval=F

---> 

<br>

----

We have modeled the dynamics of kudos tie formation. Now let's model dynamics in running behaviors.

## Covariate effects
We start with effects on behavior changes of other variables.

- the interdependence between running frequency and other sports frequency.
- gender on behavior
  
```{r echo=T, results='hide'}
myeff3 <- includeEffects(myeff2, effFrom, name = "freq_run", interaction1 = "freq_other")
myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "gender")
```
  
<br>

----
  
## Influence effects {#inf}
Last, we include effects of network position (indegree) and alter behaviors (average alter/similarity, etc.) on behavior change.
We make, for each club, 6 model specifications. We save these effect objects in a list.

- Model 1: base model + indegree effect on running
- Model 2: Model 1 + average alter effect
- Model 3: Model 1 + average attraction higher
- Model 4: Model 1 + average attraction lower
- Model 5: Model 1 + average attraction higher + lower
- Model 6: Model 1 + average similarity effect

We also fixed-and-tested the effect of outdegree on behavior, to rule out possible confounding of the outdegree effect. Score-type test indicated that outdegree-effects on behavior were 0 (not shown). 


```{r echo=T, results='hide'}
myeff0 <- myeff3                                                                            # model 0: base
myeff1 <- includeEffects(myeff0, indeg, name = "freq_run", interaction1 = "kudonet")        # model 1: indegree
myeff2 <- includeEffects(myeff1, avAlt, name = "freq_run", interaction1 = "kudonet")        # model 2: avAlt
myeff3 <- includeEffects(myeff1, avAttHigher, name = "freq_run", interaction1 = "kudonet")  # model 3: avAttHigher
myeff4 <- includeEffects(myeff1, avAttLower, name = "freq_run", interaction1 = "kudonet")   # model 4: avAttLower
myeff5 <- includeEffects(myeff3, avAttLower, name = "freq_run", interaction1 = "kudonet")   # model 5: avAttHigher+Lower
myeff6 <- includeEffects(myeff1, avSim, name = "freq_run", interaction1 = "kudonet")        # model 6: avSim

myeff <- list(myeff1, myeff2, myeff3, myeff4, myeff5, myeff6)
```

  
<br>
  
----
  
# Step 4: Estimate the model
  
Let's estimate these models. We rerun the models until adequate convergence is reached.
We store the sienaFit objects of these models in a list, which we save later on. We use 'returnDeps=TRUE' for keeping the simulated data (networks and behavior).

```{r eval= F }
m=6 # models to estimate (indeg, avAlt, avAttHigher, avAttLower, avAttHigher+Lower, avSim)

# tweak the algorithm
myalgorithm <- sienaAlgorithmCreate(projname = "test", nsub=5, n3=3 )

# siena07( myalgorithm, data = mydata, effects = myeff[[j]], prevAns= sienaFit[[j]], returnDeps=TRUE, useCluster=TRUE, nbrNodes=10, initC=TRUE, batch=TRUE)

# we make a list for storing the RSiena fit objects
sienaFit <- list()

# for club i (here, 1) we run models j in 1:m
i = 1
for (j in 1:m) {
 
# we estimate the model
try <- 1
print(paste("Estimating model ", j, " for club 1", sep=""))
sienaFit[[j]] <- siena07(myalgorithm, data = mydata, effects = myeff[[j]], returnDeps=TRUE,
                             useCluster=TRUE, nbrNodes=10, initC=TRUE, batch=TRUE) # store it in the list
    
    # re-run until we reach adequate convergence 
    while (TRUE){
      if(sienaFit[[j]]$tconv.max >= .25){
        try <- try + 1
        if (try>30) {
          print(paste("Now it lasted to long!") 
          break      
        }
        print(paste("Model did not converge adequately (", sienaFit[[j]]$tconv.max, "); ", "Repeat the estimation (", "try ", try, ")", sep = ""))
        sienaFit[[j]] <- siena07( myalgorithm, data = mydata, effects = myeff[[j]], prevAns= sienaFit[[j]], returnDeps=TRUE, useCluster=TRUE, nbrNodes=10, initC=TRUE, batch=TRUE)
      }else{
        print(paste("Reached overall maximum convergence ratio of: ", sienaFit[[j]]$tconv.max, sep = ""))
        print("")
        break
      }
    }
    
  }
  # and save the list with RSiena fit objects
  save(sienaFit, file=paste("test", "/", "sienaFit", "/", "sienaFit_club", i, ".RData", sep = ""))
  print(paste("All models are estimated for club ", i, ". Model results are stored in sienaFit_club", i, ".RData", sep=""))

}

sienaFit_clubL <- list()

for (i in 1:5) {
  temp.space <- new.env()
  bar <- load(paste("test/sienaFit/sienaFit_club", i, ".RData", sep=""), temp.space)
  sienaFit_clubL[[i]] <- get(bar, temp.space)
  rm(temp.space)
}


lapply(sienaFit_clubL, '[[', 5)
map(sienaFit_clubL, 6)


```



<br> 

Because we are now modeling the evolution of both the network and the attribute (running freq.), we will add an additional indicator to evaluate GOF; namely, does the model capture the distribution of actors’ attribute levels over time?


```{r eval=F, results='hide'}
gofbeh <- sienaGOF(sienaFit[[5]],
                   BehaviorDistribution, levls = 0:7,
                   verbose=TRUE, join=TRUE,
                   varName="freq_run")
save(gofbeh, file= paste("files", "/", "test club 1", "/", "gofbeh.RData", sep=""))
```

```{r class.source = 'fold-hide'}
load("files/test club 1/gofbeh.RData")
plot(gofbeh)
```


GOF is adequate for the distribution of running frequency values.

----

## Next up

We will check whether this model configuration also converges for the [other clubs](https://robfranken.github.io/Strava/other.html). To summarize the results over our clubs, we will perform a [meta-analysis](https://robfranken.github.io/Strava/meta.html) using a Fisher-type combination of one-tailed p-values.


----


### References
