---
title: "RSiena: Exploratory analyses "
date: "Last compiled on `r format(Sys.time(), '%B, %Y')`"
bibliography: references.bib
output:
  html_document:
    css: tweaks.css
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    code_folding: show
    code_download: yes
---

```{r, globalsettings, echo=FALSE, warning=FALSE}
library(knitr)
library(RSiena)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```



---

Let's estimate the Stochastic Actor-Oriented Model (SAOM) implemented in R as the Simulation Investigation for Empirical Network Analysis (R-SIENA), developed by @snijders2010.

----

<br>



<br> 

# Preparation

Clean the working environment.

```{r, attr.output='style="max-height: 200px;"'}
# clean the working environment 
rm (list = ls( ))
```

<br> 

We will:

1. Get our data and create an R-SIENA object
2. Inspect our data
3. Define our effects
4. Define our algorithm
5. And estimate the SAOM

Below, we will follow these steps for club 2 (N=62). We will explore the SAOM for 3 clubs in order to find the best model (see [Results](#result)). We will subsequently test if this model also fits the data of the remaining two clubs.


----

<br>

# Step 1: Data

We start off by reading in the data, specifying variable roles for the RSiena object, and creating an RSiena object. We use the clubdata.RData object.

- For now we take as our (explanatory) network variable the Kudo-network in which awarding/receiving *at least* 1 Kudo constitutes an *i,j* tie. (Ultimately, we may want to construct an ordered network-variable, in which awarding/receiving more than 1 Kudo is a sub-network of the network in which awarding/reciving 1 Kudo counts as a tie, as a robusness test) 

- Our (dependent) behavioral variable is running frequency (in times per week; ranging from 0 to 7 times per week), and running volume (in hours per week, capped at 3+).

- We included activity in other sports (e.g., cycling and swimming) as a time-varying covariate.

- And we also included gender (men vs. women and others), and winter as constant and varying covariates, respectively.


```{r}
load("clubdata.Rdata") # load (raw) club list
club <- clubdata[[2]] # grab club 

# specify variable roles for RSiena object
kudonet <- sienaDependent(club$kudo)
kudonet2 <- sienaDependent(club$kudo2) 
kudonet3 <- sienaDependent(club$kudo3) 

freq_run <- sienaDependent(club$freq_run, type= "behavior")
time_run <- sienaDependent(club$time_run, type = "behavior")
persist <- sienaDependent(club$persist, type = "behavior")

freq_other <- varCovar(club$freq_other[,,])
time_other <- varCovar(club$time_other[,,])

gender <- NA 
gender <- ifelse(club$male == 1, 1, gender)
gender <- ifelse(club$female == 1, 2, gender)
gender <- ifelse(club$other == 1, 2, gender)
gender <- coCovar(gender)

winter <- varCovar(club$winter)

#create rsiena data object
mydata <- sienaDataCreate(kudonet, freq_run, time_run, freq_other, time_other, gender, winter)
#mydata <- sienaDataCreate(kudonet, persist, freq_other, time_other, gender, winter, friendship)
```

<!--- 

After the estimation procedure, just edit the number within the double brackets [[x]] to grab another club.


For robustness check: add the number of followees as covariate
code:
following <- c(81,67,33,159,167,61,122,160,273,65,412,52,280,149,49,15,64,49,125,61,210,28,34,90,47,134,842,256,169,225,264,193,87,14,219,111,56,56,84,70,96,123,120,101,196,168,9,108,30,63,137,79,99,61,147,91,104,422,77,119,12,13)
# cluster it in 10 equally populated categories
fol <- ntile(following, 10)
following <- coCovar(fol)

---> 

----

<br>

# Step 2: Inspect data
```{r eval=F}
print01Report(mydata, modelname="files/test")
```

A text file is printed in the working directory.

![](files/test.txt){#id .class width=100% height=200px}

----

<br>

# Step 3: Define effects
We are going to define our 'myeff' object containing the model parameters. A list of all available effects for the given object can be displayed in browser by requesting effectsDocumentation(myeff). See @rsienamanual for a substantial and mathematical description of all effects.

We include:

1. [structural network effects](#str)
2. [network selection effects](#sel)
3. [network influence effects](#inf)
4. and [covariate effects](#co) on network and behavior 


```{r echo=T, results='hide'}
myeff <- getEffects(mydata)
effectsDocumentation(myeff)
```


<br>

## Structural network effects {#str}
First, we are going to include structural network effects, guided by recommendations of @snijderspres: outdegree, reciprocity, and transitivity (GWESP).

We also add degree-related effects: indegree-popularity and outdegree-activity (square-root versions). We tested whether activity and popularity effects have added value with the score-type test @Schweinberger2012 - meaning that the parameter estimates are not 0, other than the model assumes.

<!--- 

to perform score-type tests, include the effects with the setEffect command, set fix=TRUE, test=TRUE, and initialValue=0.

---> 


```{r echo=T, results='hide'}
myeff1 <- includeEffects(myeff, gwespFF, name = "kudonet") 
myeff1 <- includeEffects(myeff1, outActSqrt, inPopSqrt, name = "kudonet") 
myeff1 <- includeEffects(myeff1, isolateNet, name = "kudonet") 
```


<!--- 

test the network-isolate effect

---> 



<br>

## Selection effects {#sel}
Second, we include selection effects with respect to behavior. We estimate the effect of actors' (ego and alter) frequency, and the similarity between the two (homophily) on tie formation. Again, we applied score-type tests (not shown here).


```{r echo=T, results='hide'}
myeff2 <- includeEffects(myeff1, egoX, altX, simX, name = "kudonet", interaction1 = "freq_run")
```

<br>

## Influence effects {#inf}
We included two influence effects to test our hypotheses on the influence of social support and social comparison processes respectively:

1. The indegree effect on behavior.
This tests the social support explanation: the higher indegree (our operationalization of received support), the greater the probability of increasing the behavior. 

2. (..)


```{r echo=T, results='hide'}
myeff3 <- includeEffects(myeff2, indeg, name = "freq_run", interaction1 = "kudonet") 
myeff3 <- includeEffects(myeff3, indeg, name = "time_run", interaction1 = "kudonet")

```

<br>


<!--- 
#RF currently programming a outdegree effect on behavior, conditional on behavior alters > behavior ego.

for robustness check with separate creation and endowment effects:

myeff3 <- includeEffects(myeff2, indeg, name = "freq_run", interaction1 = "kudonet", type = "creation") 
myeff3 <- includeEffects(myeff3, indeg, name = "freq_run", interaction1 = "kudonet", type = "endow") 
myeff3 <- includeEffects(myeff3, indeg, name = "time_run", interaction1 = "kudonet", type="creation")
myeff3 <- includeEffects(myeff3, indeg, name = "time_run", interaction1 = "kudonet", type="endow")




---> 

<br>

## Covariate effects {#co}

And last we add effects on tie and behavior changes of other variables:

- the interdependence between frequency and volume
- the interdependence between running and other sports
- winter on behavior
- gender on tie selection and behavior

Note: we also estimated models with an effect of the constant dyadic covariate Friendship on Kudo ties, which was highly significant. However, this effect seemed to be highly collinear with structural network effect, meaning that it explained similar tie formation structures. Therefore we excluded Friendship. 

```{r echo=T, results='hide'}
myeff3 <- includeEffects(myeff3, effFrom, name = "time_run", interaction1 = "freq_run")
myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "time_run")

myeff3 <- includeEffects(myeff3, effFrom, name = "time_run", interaction1 = "time_other")
myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "freq_other")

myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "winter")
myeff3 <- includeEffects(myeff3, effFrom, name = "time_run", interaction1 = "winter")

myeff3 <- includeEffects( myeff3, egoX, altX, sameX, name="kudonet", interaction1 = "gender" )
myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "gender")
myeff3 <- includeEffects(myeff3, effFrom, name = "time_run", interaction1 = "gender")

```


<!--- 
#RF 
for club 4 fix the effects of gender at the current value (0), and also apply the score-type test to them.

with the following code:

myeff3 <- setEffect( myeff3, egoX, name="kudonet", interaction1 = "gender", fix=TRUE, test=TRUE, initialValue=0 )
myeff3 <- setEffect( myeff3, altX, name="kudonet", interaction1 = "gender", fix=TRUE, test=TRUE, initialValue=0 )
myeff3 <- setEffect( myeff3, simX, name="kudonet", interaction1 = "gender", fix=TRUE, test=TRUE, initialValue=0 )
myeff3 <- setEffect(myeff3, effFrom, name = "freq_run", interaction1 = "gender", fix=TRUE, test=TRUE, initialValue=0)
myeff3 <- setEffect(myeff3, effFrom, name = "time_run", interaction1 = "gender", fix=TRUE, test=TRUE, initialValue=0)


For robustness test with individual level covariate 'following', include effect on 
behavior and ties: 

myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "following")
myeff3 <- includeEffects( myeff3, egoX, name="kudonet", interaction1 = "following" )

---> 



<br>

Now check which effects are included in the myeff object.

```{r class.source = 'fold-hide' }
options(width = 100) # ignore (this is for the html formatting)
print(myeff3)
```


Seems allright!

----

<br>

# Step 4: Define algorithm

```{r class.source = 'fold-hide'  }
myalgorithm <- sienaAlgorithmCreate(projname = "test")
```

----

<br>

# Step 5: Estimate the model

Let's estimate the SAOM.

We will save the results in a html-table using the siena.table()-function, in the 'files'-folder in our working directory. We will run the model as many times as necessary, until we reach a convergence ratio of < .25. We use 'returnDeps=TRUE' for keeping the simulated data (networks and behavior)


```{r eval= F }
try <- 1

ansM1 <- siena07(myalgorithm, data = mydata, effects = myeff3, returnDeps=TRUE)
siena.table(ansM1, type="html", tstat=T, d=2, sig=T, file = paste("files", "/", "last.html", sep = ""))

# the following script lets the model re-run until we get a good convergence ratio
while (TRUE){
  
  if(ansM1$tconv.max >= .25){
    try <- try + 1
    print(paste("Try:", try, sep=" "))
    ansM1 <- siena07( myalgorithm, data = mydata, effects = myeff3, prevAns= ansM1, returnDeps=TRUE)
    siena.table(ansM1, type="html", tstat=T, d=2, sig=T, file = paste("files", "/", "ansM1_", try, ".html", sep=""))
    
  }else{
    siena.table(ansM1, type="html", tstat=T, d=2, sig=T, file = paste("files", "/", "final", ".html", sep=""))
    print(paste("Reached convergence ratio of ", ansM1$tconv.max, sep = ""))
    break
  }
}
```

<br>


# Step 6: Goodness-of-fit

First define some functions for GOF (see [here](http://www.stats.ox.ac.uk/~snijders/siena/sienaGOF_vdB.R)).

```{r class.source = 'fold-hide'}
# GeodesicDistribution calculates the distribution of non-directed
   # geodesic distances; see ?sna::geodist
GeodesicDistribution <- function (i, data, sims, period, groupName,
   varName, levls=c(1:5,Inf), cumulative=TRUE, ...) {
     x <- networkExtraction(i, data, sims, period, groupName, varName)
     require(sna)
     a <- sna::geodist(symmetrize(x))$gdist
     if (cumulative)
     {
       gdi <- sapply(levls, function(i){ sum(a<=i) })
     }
     else
     {
       gdi <- sapply(levls, function(i){ sum(a==i) })
     }
     names(gdi) <- as.character(levls)
     gdi
}

# The following function is taken from the help page for sienaTest
testall <- function(ans){
    for (i in which(ans$test)){
    sct <- score.Test(ans,i)
    cat(ans$requestedEffects$effectName[i], '\n')
    print(sct)}
    invisible(score.Test(ans))
}
```

Now we can apply sienaGOF to the data

```{r eval= F}
# indegree
gofi0 <- sienaGOF(ansM1, IndegreeDistribution, cumulative=FALSE,
     verbose=TRUE, join=TRUE, varName="kudonet")
plot(gofi0)
# outdegree
gofo0 <- sienaGOF(ansM1, OutdegreeDistribution, cumulative=FALSE,
     verbose=TRUE, join=TRUE, varName="kudonet")
plot(gofo0)
# triad census
gof0.tc <- sienaGOF(ansM1, TriadCensus,
     verbose=TRUE, join=TRUE, varName="kudonet")
plot(gof0.tc)
# geodesic distances
gof0.gd <- sienaGOF(ansM1, GeodesicDistribution, levls=c(1:5), cumulative=FALSE,
     verbose=TRUE, join=TRUE, varName="kudonet")
plot(gof0.gd)

```



<!--- 

#RF in the files folder sub-folders are made for each club, with preliminary results (final.html) in the 'prelim' folders


---> 

----

<br>

# Step 7: Results {#result}

We present the SAOM results for all clubs.

## {.tabset .tabset-fade}

### Club 1
```{r, echo=F}
htmltools::includeHTML("files/prelim3/club 1/final.html")

```

### Club 2
```{r, echo=F}
htmltools::includeHTML("files/prelim3/club 2/final.html")
```

### Club 3
```{r, echo=F}
htmltools::includeHTML("files/prelim3/club 3/final.html")
```

### Club 4
```{r, echo=F}
htmltools::includeHTML("files/prelim3/club 4/final.html")
```

Note that we excluded gender effects on tie formation and behavior change, as they were non-significant and caused convergence issues (probably because of small N). Also note the high rate parameters!

### Club 5
```{r, echo=F}
htmltools::includeHTML("files/prelim3/club 5/final.html")
```

## {-}

- We see a marginal reciprocated degree effect on running volume (club 1) and running frequency (club 2).

- We see positive average reciprocated alter effect on running frequency and volume (club 5). 


Thus, in three clubs we find influence effects (indicating support and comparison). These effects are observed even after controlling for selection and structural network effects. 



----

## Ordered network variable {#lvl2}

In the estimated SAOM the reciprocation of at least 1 Kudo constituted the presence of a tie. I also estimated the model with reciprocation of *more than* 1 Kudo constituting the presence of a tie

<!--- 

replace the kudonet with kudonet2
---> 

### {.tabset .tabset-fade}

#### Club 1
```{r, echo=F}
htmltools::includeHTML("files/clubs results/knet2/club1.html")
```

#### Club 2
```{r, echo=F}
htmltools::includeHTML("files/clubs results/knet2/club2.html")
```

#### Club 5
```{r, echo=F}
htmltools::includeHTML("files/clubs results/knet2/club5.html")
```

### {-}



----

<br>


### References
