---
title: "Strava paper"
bibliography: references.bib
date: "Last compiled on `r format(Sys.time(), '%B, %Y')`"
output: 
  html_document:
    css: tweaks.css
    toc:  false
    toc_float: true
    number_sections: false
    code_folding: show
    code_download: yes
---

```{r, globalsettings, echo=FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, results="hold", message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

----  


We will show how the influence statistics in RSiena behave. 
We assume that ego is connected to all alters. 

<br>

## Functions

First, we define some functions:

<br>

### Centering behavior variables
```{r}
fcentering <- function(actors){
centered <- actors - mean(actors)
return(centered)
}
```

<br> 

### Calculating similarity scores

$$ sim^z_{ij} = 1 - \frac{|z_i - z_j|}{r_V} $$
```{r}
fsimij <- function(actors, min, max){
#rv <- max(actors) - min(actors)
rv <- max - min
mat1 <- matrix(actors, nrow=length(actors), ncol=length(actors), byrow=TRUE)
mat2 <- t(mat1)
simij <- 1 - ( abs(mat1-mat2) / rv)
return(simij)
}
```

<br> 


### The average similarity effect (avSim)

$$ s^{beh}_{i5}(x,z) = x^{-1}_{i+}\sum_j x_{ij}(sim^z_{ij} - \widehat{sim^z}) $$
```{r}
favSim <- function(ego, alters, min, max) {
  actors <- c(ego,alters) #define the network
  beh_centered <- fcentering(actors) #center behavior scores
  simij <- fsimij(beh_centered, min, max) #calculate the similarity scores
  diag(simij) <- NA
  msimij <- mean(simij, na.rm=TRUE) #calculate the mean similarity score. only calculate mean on non-diagonal cells??!!
  simij_c <- simij - msimij #center the similarity scores
  
  statistic <- sum(simij_c[1,], na.rm = TRUE) / length(alters) #the actual statistic
  
  return(statistic)
}
```
```{r}
favSim2 <- function(ego, alters, min, max) {
  actors <- c(ego,alters) #define the network
  beh_centered <- fcentring(actors) #center behavior scores
  simij <- fsimij(beh_centered, min, max) #calculate the similarity scores
  diag(simij) <- NA
  #msimij <- mean(simij, na.rm=TRUE) #calculate the mean similarity score. only calculate mean on non-diagonal cells??!!
  simij_c <- simij # - msimij #center the similarity scores
  
  statistic <- sum(simij_c[1,], na.rm = TRUE) / length(alters) #the actual statistic
  
  return(statistic)
}
```

<br> 


### The average attraction towards higher effect (avAttHigher)

$$ s^{beh}_{i23}(x,z) = x^{-1}_{i+}\sum_j x_{ij} (\, I \{ (z_j)>z_i \} sim^z_{ij} + I \{ (z_j ) \le z_i \} )\,$$
```{r}
favAttHigher <- function(ego, alters, min, max) {
  actors <- c(ego,alters)
  beh_centered <- fcentering(actors)
  simij <- fsimij(beh_centered, min, max)
  diag(simij) <- NA
  
  simijH <- simij[1,]
  simijH[beh_centered <= beh_centered[1]] <- 1
  simijH[1] <- NA
  statistic <- sum(simijH, na.rm = TRUE) / length(alters)
  
 
  return(statistic)
}
```

<br> 

### The average attraction towards higher alters effect (avAttLower)

$$ s^{beh}_{i24}(x,z) = x^{-1}_{i+}\sum_j x_{ij} (\, I \{ (z_j)<z_i \} sim^z_{ij} + I \{ (z_j ) \ge z_i \} )\,$$
```{r}
favAttLower <- function(ego, alters, min, max) {
  actors <- c(ego,alters)
  beh_centered <- fcentering(actors)
  simij <- fsimij(beh_centered, min, max)
  diag(simij) <- NA
  
  simijL <- simij[1,]
  simijL[beh_centered >= beh_centered[1]] <- 1
  simijL[1] <- NA
  statistic <- sum(simijL, na.rm = TRUE) / length(alters)
  
 
  return(statistic)
}
```

<br> 


### Average alter effect (avAlt)

$$ s^{beh}_{i27}(x,z) = z_i(\textstyle \sum_jx_{ij}z_j)/(\textstyle \sum_jx_{ij}) $$ 

```{r}
favAlt <- function(ego, alters, ...) {
  actors <- c(ego,alters)
  beh_centered <- fcentering(actors)
  
  statistic <- beh_centered[1] * (sum(beh_centered[-1], na.rm = TRUE) / length(alters))
  
 
  return(statistic)
}
```

<br> 


### Average attraction mean (AttMean)

We are attracted to the mean of our alters! 

(behaves approximately similar to avAlt, different shape.)

$$ s^{beh}_{i}(x,z) = 1 - \frac{\lvert(z_i - x^{-1}_{i+}\sum_j x_{ij}z_j)\rvert}{r_Z}  $$

```{r}
fAttMean <- function(ego, alters, min, max, ...) {
  rv <- max - min
  actors <- c(ego,alters)
  beh_centered <- fcentring(actors)
  
  statistic <- 1 -  abs(beh_centered[1] - (sum(beh_centered[-1], na.rm = TRUE) / length(alters)))/rv #thus we strive for a highest local similarity score!
  
 
  return(statistic)
}
```

<br>

## Make plot of statistic values {.tabset .tabset-fade}

```{r}

finluenceplot <- function(alters, min, max, fun, params, results=TRUE, plot=TRUE) {
  #check correct number of parameters are given
   if (length(fun) != length(params)) stop("Please provide one (and only one) parameter for each of the behavioral effects!")
  
  #calculuate value of evaluation function
  s <- NA
  for (i in min:max) {
    s[i] <- 0
    for (j in 1:length(fun)) {
      s[i] <- s[i] + params[j]*fun[[j]](i, alters, min, max)      
    }
  }
  
  #calculate the probabilities  
  p <- NA
  for (i in min:max) {
    p[i] <- exp(s[i]) / sum(exp(s))
  }
  
  #calculate the probabilities of choice set  
  p2 <- NA
  for (i in min:max) {
    if (i==min) { 
      p2[i] <- exp(s[i]) / sum(exp(s[i]) + exp(s[i + 1])) 
    } else if (i==max) { 
      p2[i] <- exp(s[i]) / sum(exp(s[i]) + exp(s[i - 1]))
    } else {
      p2[i] <- exp(s[i]) / sum(exp(s[i]) + exp(s[i - 1]) + exp(s[i + 1]))
    }
  }
  
  #calculate the probability ratio  
  r <- NA
  for (i in min:max) {
    r[i] <- p[i] / p[1]
  }
  
  #some simple plots
  if (plot) { 
      name <- deparse(substitute(fun))
      name <- stringr::str_sub(as.character(name), 6, -2)
      par(mfrow=c(2,2))
      plot(y=s, x=min:max, xlab="ego behavioral score", ylab=name, type="b")
      mtext("EVALUATION", line=1)
      mtext(paste("alters:", paste0(alters, collapse=",")))
      plot(y=p, x=min:max, xlab="ego behavioral score", ylab=name, ylim=c(0,1), type="b")
      mtext("PROBABILITIES", line=1)
      mtext(paste("alters:", paste0(alters, collapse=",")))
      plot(y=p2, x=min:max, xlab="ego behavioral score", ylab=name, ylim=c(0,1), type="b")
      mtext("PROBABILITIES_choice_set", line=1)
      mtext(paste("alters:", paste0(alters, collapse=",")))
      plot(y=r, x=min:max, xlab="ego behavioral score", ylab=name, type="b")
      mtext("RATIOS", line=1)
      mtext(paste("alters:", paste0(alters, collapse=",")))
  }
  
  #return results for more fancy plotting
  if (results) {
    x <- min:max
    df <- data.frame(x, s, p,p2, r)
    return(df)
  }
  
}
```

<br>

## Examples

```{r}
alters1 <- rep(c(3,3,3,3,3,3),4)

finluenceplot(alters=alters1, min=1, max=6, list(flinear, fquad), params=c(.1,-1))

finluenceplot(alters=alters1, min=1, max=6, list(favSim), params=c(1))
finluenceplot(alters=alters1, min=1, max=6, list(favAlt), params=c(1))
finluenceplot(alters=alters1, min=1, max=6, list(fAttMean), params=c(1))
finluenceplot(alters=alters1, min=1, max=6, list(favAttHigher), params=c(1))
finluenceplot(alters=alters1, min=1, max=6, list(favAttLower), params=c(1))
finluenceplot(alters=alters1, min=1, max=6, fun=list(favAttLower, favAttHigher), params=c(1,1))
finluenceplot(alters=alters1, min=1, max=6, list(fAttMean, favAttHigher), params=c(1,1))


```

```{r}
alters1 <- rep(c(1,1,1,6,6,6), 40)
finluenceplot(alters=alters1, min=1, max=6, list(favSim), params=c(1))
finluenceplot(alters=alters1, min=1, max=6, list(favAlt), params=c(1))
finluenceplot(alters=alters1, min=1, max=6, list(fAttMean), params=c(1))
finluenceplot(alters=alters1, min=1, max=6, fun=list(favAttLower, favAttHigher), params=c(1,0))
finluenceplot(alters=alters1, min=1, max=6, fun=list(favAttLower, favAttHigher), params=c(0,1))

finluenceplot(alters=alters1, min=1, max=6, fun=list(favAttLower, favAttHigher), params=c(1,3))

finluenceplot(alters=alters1, min=1, max=6, fun=list(favAttLower, favAttHigher), params=c(1,-1))
finluenceplot(alters=alters1, min=1, max=6, fun=list(favAttLower, favAttHigher), params=c(-1,1))
finluenceplot(alters=alters1, min=1, max=6, fun=list(flinear, favAttLower, favAttHigher), params=c(1,-1,1))
finluenceplot(alters=alters1, min=1, max=6, fun=list(flinear, favAttLower, favAttHigher), params=c(1,-1,1))
finluenceplot(alters=alters1, min=1, max=6, fun=list(flinear), params=c(1))

```


```{r}

alters1 <- rep(c(1,2,2), 100)
finluenceplot(alters=alters1, min=1, max=6, list(favSim), params=c(1))
finluenceplot(alters=alters1, min=1, max=6, list(favAlt), params=c(1))
finluenceplot(alters=alters1, min=1, max=6, list(fAttMean), params=c(1))
finluenceplot(alters=alters1, min=1, max=6, fun=list(favAttLower, favAttHigher), params=c(1,0))
finluenceplot(alters=alters1, min=1, max=6, fun=list(favAttLower, favAttHigher), params=c(0,1))
finluenceplot(alters=alters1, min=1, max=6, fun=list(favAttLower, favAttHigher), params=c(1,1))
finluenceplot(alters=alters1, min=1, max=6, fun=list(favAttLower, favAttHigher), params=c(1,-1))
finluenceplot(alters=alters1, min=1, max=6, fun=list(favAttLower, favAttHigher), params=c(-1,1))

```

<br>

## Results model strava

```{r}
alters1 <- rep(c(2, 2, 2, 5, 5, 5))
alters2 <- rep(c(3.5, 3.5, 3.5, 3.5, 3.5, 3.5))
``` 

```{r}
# club1
finluenceplot(alters = alters1, min = 1, max = 6, results = FALSE, fun = list(flinear, fquad, favAttHigher,
    favAttLower), params = c(-0.15, -0.002, 2.0027, 6.0905))


finluenceplot(alters = alters2, min = 1, max = 6, results = FALSE, fun = list(flinear, fquad, favAttHigher,
    favAttLower), params = c(-0.15, -0.002, 2.0027, 6.0905))

# club2
finluenceplot(alters = alters1, min = 1, max = 6, results = FALSE, fun = list(flinear, fquad, favAttHigher,
    favAttLower), params = c(-0.24, 0.0075, 3.3213, 5.0047))


finluenceplot(alters = alters2, min = 1, max = 6, results = FALSE, fun = list(flinear, fquad, favAttHigher,
    favAttLower), params = c(-0.24, 0.0075, 3.3213, 5.0047))


# club3
finluenceplot(alters = alters1, min = 1, max = 6, results = FALSE, fun = list(flinear, fquad, favAttHigher,
    favAttLower), params = c(-0.2085, 0.021, 3.7703, 2.1283))


finluenceplot(alters = alters2, min = 1, max = 6, results = FALSE, fun = list(flinear, fquad, favAttHigher,
    favAttLower), params = c(-0.2085, 0.021, 3.7703, 2.1283))


# club4
finluenceplot(alters = alters1, min = 1, max = 6, results = FALSE, fun = list(flinear, fquad, favAttHigher,
    favAttLower), params = c(-1.3313, 0.0192, 2.5493, 4.4952))


finluenceplot(alters = alters2, min = 1, max = 6, results = FALSE, fun = list(flinear, fquad, favAttHigher,
    favAttLower), params = c(-1.3313, 0.0192, 2.5493, 4.4952))


# club5
finluenceplot(alters = alters1, min = 1, max = 6, results = FALSE, fun = list(flinear, fquad, favAttHigher,
    favAttLower), params = c(0.1021, 0.0743, 1.1736, 8.5486))


finluenceplot(alters = alters2, min = 1, max = 6, results = FALSE, fun = list(flinear, fquad, favAttHigher,
    favAttLower), params = c(0.1021, 0.0743, 1.1736, 8.5486))
```
