---
title: "Meta-analysis"
date: "Last compiled on `r format(Sys.time(), '%B, %Y')`"
bibliography: references.bib
output:
  html_document:
    css: tweaks.css
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    code_folding: show
    code_download: yes
---

```{r, globalsettings, echo=FALSE, warning=FALSE}
library(knitr)
library(RSiena)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```


---

We also built increasingly complex models for the remaining 4 clubs. We started by modeling network structure and dynamics in kudos tie formation, using a rather simple model specification. If this did not produce adequate GOF, we assessed how the model may be extended, by a theory-guided exploration. We performed a forward model-selection procedure and use score-type tests to see which effect could improve the model. We use *sienaGOF* to estimate how the fit (based on Mahalanobis distance) would increase had we included the fixed-and-tested effects (hence, circumventing the need for re-estimating the model with the new effect). This resulted in slightly different model-specification for kudos-tie formation dynamics for each club.

If adequate GOF was produced, we started modeling running activity dynamics, and investigated how well our model was able to reproduce the distribution of running behavior values of our panel data. 

Down below, we create the final effect-objects for each club. 
Note: for subsequent meta-analysis, we needed to ensure that the model specification for all our club-networks is identical. In some clubs, some effects were rather important; in clubs where this was not the case, we fixed these to 0.


# Preparation
Clean the working environment and read in the R-SIENA object list.

```{r, attr.output='style="max-height: 200px;"'}
# clean the working environment 
rm (list = ls( ))
# load in the R-SIENA objects
load("clubdata_rsiena.Rdata")
```

# Effect objects {.tabset .tabset-fade}

## Club 2

## Club 3

## Club 4

## Club 5

# {-}

----

<br>

# Results {.tabset .tabset-fade}


We estimated the models for each club:

- Model 1: the baseline model + indegree effect
- Model 2: we add the average attraction higher and lower effects.

We altered the algorithm with the following parameters: *nsub*=5, *n3*=5000 (see @rsienamanual, 6.3.3).

Down below, we present the results:


## Club 2 {.tabset .tabset-fade}

### Model 1

### Model 2

## {-}

## Club 3 {.tabset .tabset-fade}

### Model 1

### Model 2

## {-}

## Club 4 {.tabset .tabset-fade}

### Model 1

### Model 2

## {-}

## Club 5 {.tabset .tabset-fade}

### Model 1

### Model 2

## {-}

# {-}

----

<br>

# Meta-analysis 

To summarize the results over our clubs, we will perform a meta-analysis using a Fisher-type combination of one-tailed p-values, described by @hedges2014 and @bosker2011 (Chapter 3). 

We perform a double test:

1. For detecting if there are any club-networks with a positive parameter value, the null hypothesis tested is: for all clubs, the value of this parameter is zero or less than zero; with the alternative hypothesis: for at least one club, the value of this parameter is greater than zero.
2. For detecting if there are any club-networks with a negative parameter value, the null hypothesis tested is: for all clubs, the value of this parameter is zero or greater than zero; with the alternative hypothesis: for at least one club, the value of this parameter is less than zero.

For each of these combined tests, the *p*-value is given. We use for each test the significance level of *α*/2, which yields an overall combined test at significance level *α*. Four different overall results are possible. Indicating the right-sided and left-sided *p*-values by *p<sub>r</sub>* and *p<sub>l</sub>* respectively, these results are:

- *p<sub>r</sub>* $\gt$ *α*/2 & *p<sub>l</sub>* $\gt$ *α*/2:

No evidence for any non-zero parameter values.

- *p<sub>r</sub>* $\leq$ *α*/2 & *p<sub>l</sub>* $\gt$ *α*/2:

Evidence that some club-networks have a positive parameter value, but no evidence for any negative parameter values.

- *p<sub>r</sub>* $\gt$ *α*/2 & *p<sub>l</sub>* $\leq$ *α*/2:

Evidence that some clubs have a negative parameter value, no evidence for any positive parameter values

- *p<sub>r</sub>* $\leq$ *α*/2 & *p<sub>l</sub>* $\leq$ *α*/2:

Evidence that some clubs have a negative parameter value, and some others have a positive parameter value.

<br>

----


<br>

----


### References 
