---
title: "Meta-analysis"
date: "Last compiled on `r format(Sys.time(), '%B, %Y')`"
bibliography: references.bib
output:
  html_document:
    css: tweaks.css
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    code_folding: show
    code_download: yes
---

```{r, globalsettings, echo=FALSE, warning=FALSE}
library(knitr)
library(RSiena)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```


---

We will estimate the SAOM for the remaining 4 clubs.

<br>

# Preparation
Clean the working environment and read in the RSiena object list.

```{r, attr.output='style="max-height: 200px;"'}
# clean the working environment 
rm (list = ls( ))
# load in the R-SIENA objects
load("clubdata_rsiena.Rdata")
```

<br>

----

# Estimate SAOMs for other clubs
For each club, we incrementally build the model and assess GOF.


## {.tabset .tabset-fade}

### Club 2

First, we model network structure and dynamics in kudos tie formation. 

We include the following structural effects: *outdegree*, *reciprocity*, *transitivity*. Also, we added degree-related endogenous effects: *outdegree-activity* and *indegree-popularity*.

In this club, the *outIsolate*-effect (leading to now giving kudos to anyone) was rather important: a score-type test indicated that it was not 0 (not shown).

This rather simple model specification did not produce sufficient GOF.
GOF based on indegree-distribution is good, but the fit based on outdegree- (p=0.001) and geodesic distance-distribution (p=0) was poor (not shown). 

We investigate how the model may be extended to obtain a better fit. We do this by a theory-guided exploration. We include the following effects: 

- effects that model the inverse indication of local hierarchy (*3-cycle* and *reciprocity x GWESP*)
- the *reciprocal degree-related activity* effect (reciAct; for which we expect a negative parameter estimate)
- an effect to model the relationsip between in- and outdegrees (*outdegree-popularity*)

Based on this, we do a foward model-selection and use the score-type test to see which of these effects could improve the model (spoiler: all of them do!). 

We use *sienaGOF* to estimate how the fit (based on Mahalanobis distance) would increase had we included the fixed-and-tested effects (hence, circumventing the need for re-estimating the model with the new effect). 

<!--- 

@RF: 
in all networks, out- and indegrees are strongly dispersed. So we may also want to let the rate function condition on the log-outdegree (outRateLog)

---> 
 

```{r echo=T, results='hide'}
mydata <- clubdata_rsiena[[2]] # grab club

myeff <- getEffects(mydata) # define myeff object

# structural network effects
myeff1 <- includeEffects(myeff, gwespFF, name = "kudonet") 
myeff1 <- includeEffects(myeff1, outActSqrt, inPopSqrt, name = "kudonet") 
myeff1 <- includeEffects( myeff1, outIso, name = "kudonet")

# we fix-and-test these effects:

# 1. interaction GWESP x recip.
myeff1 <- includeInteraction(myeff1, recip, gwespFF, parameter = 69, name = "kudonet")
myeff1 # the interaction-effect is 18
eff1 <- myeff1[myeff1$include,]$effect1[18]
eff2 <- myeff1[myeff1$include,]$effect2[18]
myeff1 <- setEffect(myeff1, unspInt, fix=TRUE, test=TRUE,
                                    effect1=eff1, effect2=eff2)
# 2. 3-cycles effect
myeff1 <- setEffect(myeff1, cycle3, name = "kudonet", fix=TRUE, test=TRUE, include=TRUE)
# 3. reciprocal degree-related activity effect
myeff1 <- setEffect(myeff1, reciAct, name = "kudonet", fix=TRUE, test=TRUE, include=TRUE)
# 4. outdegree popularity effect
myeff1 <- setEffect( myeff1, outPopSqrt, name = "kudonet", fix=TRUE, test=TRUE, include=TRUE)

# selection effects for kudos ties formation based on running activity
myeff2 <- setEffect(myeff1, higher, name = "kudonet", interaction1 = "freq_run")

# and selection based on gender 
myeff2 <- includeEffects(myeff2, egoX, altX, sameX, name="kudonet", interaction1 = "gender" )
```

<br>

We (1) define the algorithm, (2) estimate the model, and (3) assess GOF based on three auxiliary network statistics: outdegree, indegree and geodesic distance distribution. We investigate how GOF based on these diagnostics would improve had we included the fixed-and-tested effects. Note that these effects are not controlled for each other, only for the estimated parameters. 

```{r eval=F, results='hide'}
myalgorithm <- sienaAlgorithmCreate(projname = "test") # set the SAOM algorithm 

ansM1 <- siena07(myalgorithm, data = mydata, effects = myeff2, # estimate the SAOM
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE)

# we can check the score-test over the 4-degree of freedom aggregate.
# and seperately, by calling:
summary(ansM1)

# GOF diagnostics:
gofi <- sienaGOF(ansM1, 
                 IndegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

gofo <- sienaGOF(ansM1, 
                 OutdegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

GeodesicDistribution <- function (i, data, sims, period, groupName,
   varName, levls=c(1:5, Inf), cumulative=TRUE, ...) {
     x <- networkExtraction(i, data, sims, period, groupName, varName)
     require(sna)
     a <- sna::geodist(symmetrize(x))$gdist
     if (cumulative)
     {
       gdi <- sapply(levls, function(i){ sum(a<=i) })
     }
     else
     {
       gdi <- sapply(levls, function(i){ sum(a==i) })
     }
     names(gdi) <- as.character(levls)
     gdi
}

gofgeo <- sienaGOF(ansM1, 
                 GeodesicDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

goflist <- list(gofi, gofo, gofgeo)
save(goflist, file= paste("files", "/", "other clubs", "/", "club 2", "/", "gof1.RData", sep=""))
```

#### {.tabset .tabset-fade}

##### Indegree distribution
```{r class.source = 'fold-hide'}
load("files/other clubs/club 2/gof1.RData")
plot(goflist[[1]])
```

Indegree GOF is still all right.

<br>

----

##### Outdegree distribution
```{r class.source = 'fold-hide'}
plot(goflist[[2]])
summary(goflist[[2]])
```

<br>

Outdegree GOF is bad (but not p=0). MHD is the predicted Mahalanobis distance in the extended model (including the particular effect). We see that the Outdegree-distribution would improve drastically if we include the interaction between reciprocity and transitivity (GWESP).

<br>

----

##### Geodesic distance distribution
```{r class.source = 'fold-hide'}
plot(goflist[[3]])
summary(goflist[[3]])
```

<br>

GD-distribution fit would improve most after including the reciAct-effect, closely followed by the recip x GWESP-effect. Since the former effect does not improve the MD for the outdegree-distribution, inclusion of the recip x GWESP-effect seems most promising.  

<br>

----


#### {-}

We re-estimate the model with with the additional reciprocity-GWESP interaction-effect. We use the prevAns-option to use previous model-results as starting values for the estimation.

```{r eval=F, results='hide'}
myeff2 <- includeInteraction(myeff2, recip, gwespFF, parameter = 69, name = "kudonet")

ansM1 <- siena07(myalgorithm, data = mydata, effects = myeff2, prevAns = ansM1, # prior estimates as starting values
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE)

# GOF diagnostics:
gofi <- sienaGOF(ansM1, 
                 IndegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

gofo <- sienaGOF(ansM1, 
                 OutdegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

gofgeo <- sienaGOF(ansM1, 
                 GeodesicDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

goflist <- list(gofi, gofo, gofgeo)
save(goflist, file= paste("files", "/", "other clubs", "/", "club 2", "/", "gof1.2.RData", sep=""))
```

#### {.tabset .tabset-fade}

##### Indegree distribution
```{r class.source = 'fold-hide'}
load("files/other clubs/club 2/gof1.2.RData")
plot(goflist[[1]])
```

Indegree GOF is still all right.

<br>

----

##### Outdegree distribution
```{r class.source = 'fold-hide'}
plot(goflist[[2]])
```

The fit for the outdegree distribution is now good!

<br>

----

##### Geodesic distance distribution
```{r class.source = 'fold-hide'}
summary(goflist[[3]])
```

<br>

Geodesic distance distribution MD is still poor, although the p=0 has disapeared. The GD-fit is predicted to improve a lot after including the reciAct effect. 

<br>

----


#### {-}

Let's  include the reciAct effect, to see how this improves geodesic distance GOF. We re-estimate the model with the new effect.

```{r eval=F, results='hide'}
myeff2 <- includeEffects(myeff2, reciAct, name = "kudonet")

ansM1 <- siena07(myalgorithm, data = mydata, effects = myeff2, prevAns = ansM1, # prior estimates as starting values
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE)

# GOF diagnostics:
gofi <- sienaGOF(ansM1, 
                 IndegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

gofo <- sienaGOF(ansM1, 
                 OutdegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

gofgeo <- sienaGOF(ansM1, 
                 GeodesicDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

goflist <- list(gofi, gofo, gofgeo)
save(goflist, file= paste("files", "/", "other clubs", "/", "club 2", "/", "gof1.3.RData", sep=""))
```

#### {.tabset .tabset-fade}

##### Indegree distribution
```{r class.source = 'fold-hide'}
load("files/other clubs/club 2/gof1.3.RData")
plot(goflist[[1]])
```

Indegree GOF is still all right.

<br>

----

##### Outdegree distribution
```{r class.source = 'fold-hide'}
plot(goflist[[2]])
```

Outdegree GOF is still all right.

<br>

----

##### Geodesic distance distribution
```{r class.source = 'fold-hide'}
plot(goflist[[3]])
```

The fit for the GD distribution is now good!

<br>

We fine-tune the network-part of the model: we include some parameters as fixed to 0 and we exclude 3-cycles, for subsequent meta-analysis:

```{r echo=T, results='hide'}
myeff2 <- setEffect(myeff2, cycle3, name = "kudonet", include=FALSE)
myeff2 <- setEffect(myeff2, outPopSqrt, name = "kudonet", fix=TRUE, test=FALSE, initialValue = 0)
```

<br>

----

#### {-}


Now that we reached adequate GOF for the kudos-network, it is time to model running activity dynamics:

- interdepence between frequency and duration of running.
- interdependence between running and other sports. We seem to have a problem of overfitting. We exlude the effects of other sports (the score-type test indicates that they are 0).
- gender-effects on behavior change are also fixed-and-tested: they are not different than 0.
- two influence effects: *indegree* and *upward assimilation*. We control for the outdegree-effect: we fix it and score-type tested it.

```{r echo=T, results='hide'}
myeff3 <- includeEffects(myeff2, effFrom, name = "time_run", interaction1 = "freq_run")
myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "time_run")
myeff3 <- includeEffects(myeff3, effFrom, name = "time_run", interaction1 = "time_other")
myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "freq_other")

# gender effects fixed-and-tested
myeff3 <- setEffect(myeff3, effFrom, name = "freq_run", interaction1 = "gender", fix=T, test = F, initialValue = 0)
myeff3 <- setEffect(myeff3, effFrom, name = "time_run", interaction1 = "gender", fix=T, test = F, initialValue = 0)
# same for 'other sports'
myeff3 <- setEffect(myeff3, effFrom, name = "time_run", interaction1 = "time_other", fix=T, test=F, initialValue=0)
myeff3 <- setEffect(myeff3, effFrom, name = "freq_run", interaction1 = "freq_other", fix=T, test=F, initialValue=0)

# influence effects:
# 1. indegree effect
myeff3 <- includeEffects(myeff3, indeg, name = "freq_run", interaction1 = "kudonet") 
myeff3 <- includeEffects(myeff3, indeg, name = "time_run", interaction1 = "kudonet")
# 2. upward assimilation
myeff4 <- includeEffects(myeff3, avAttHigher, name = "freq_run", interaction1 = "kudonet") 
myeff4 <- includeEffects(myeff4, avAttHigher, name = "time_run", interaction1 = "kudonet")

# outdegree effect (fixed-and-tested)
myeff4 <- setEffect(myeff4, outdeg, name = "freq_run", interaction1 = "kudonet", fix=T, test=F, initialValue=0)
myeff4 <- setEffect(myeff4, outdeg, name = "time_run", interaction1 = "kudonet", fix=T, test=F, initialValue=0)
```


<br> 

We estimate the model and investigate another GOF-diagnostic: the distribution of running behavior variables. We estimate (1) a model with the objective function for running activity comprising the basic shape effects, covariate effects, and *indegree* effect; and (2) a model including the *upward assimilation* and *outdegree* effect. 

We run these models until we reach convergence. We save the sienaFit-objects of both models for subsequent meta-analyses.

#### {.tabset .tabset-fade}

##### The indegree-effect
```{r eval=F, results='hide'}
try <- 1
ansM2 <- siena07(myalgorithm, data = mydata, effects = myeff3, # estimate the SAOM
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE, prevAns=ansM1)

# the following script lets the model re-run until we get a good convergence ratio
while (TRUE){
  
  if(ansM2$tconv.max >= .25){
    try <- try + 1
    print(paste("Model did not converge: ", ansM2$tconv.max, sep = ""))
    print(paste("Try:", try, sep=" "))
    ansM2 <- siena07( myalgorithm, data = mydata, effects = myeff3, prevAns= ansM2, returnDeps=TRUE)
    
  }else{
    siena.table(ansM2, type="html", tstat=T, d=3, sig=T, file = paste("files", "/", "other clubs", "/", "club 2", "/", "Model 2", ".html", sep=""))
    print(paste("Reached convergence ratio of ", ansM2$tconv.max, sep = ""))
    break
  }
}

gofbeh <- sienaGOF(ansM2,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "freq_run")

gofbeh2 <- sienaGOF(ansM2,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "time_run")

goflist2 <- list(gofbeh, gofbeh2)
save(goflist2, file= paste("files", "/", "other clubs", "/", "club 2", "/", "gof2.RData", sep=""))

# save the sienaFit object
save(ansM2, file= paste("files", "/", "other clubs", "/", "club 2", "/", "model2.RData", sep=""))
```

<br>

**Violin plot of running frequency**
```{r class.source = 'fold-hide'}
load("files/other clubs/club 2/gof2.RData")
plot(goflist2[[1]])
```

**Violin plot of running duration**
```{r class.source = 'fold-hide'}
plot(goflist2[[2]])
```

Acceptable GOF for frequency, not for duration...

##### The upward assimilation effect

We excluded (and tested) effects on behavior change of other sports activities. These seemed to cause convergence-issues (overfitting), but they were 0.

```{r eval=F, results='hide'}
try <- 1
ansM3 <- siena07(myalgorithm, data = mydata, effects = myeff4, # estimate the SAOM
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE, prevAns=ansM2)

# the following script lets the model re-run until we get a good convergence ratio
while (TRUE){
  
  if(ansM3$tconv.max >= .25){
    try <- try + 1
    print(paste("Model did not converge: ", ansM3$tconv.max, sep = ""))
    print(paste("Try:", try, sep=" "))
    ansM3 <- siena07( myalgorithm, data = mydata, effects = myeff4, prevAns= ansM3, returnDeps=TRUE)
    
  }else{
    siena.table(ansM3, type="html", tstat=T, d=3, sig=T, file = paste("files", "/", "other clubs", "/", "club 4", "/", "Model 3", ".html", sep=""))
    print(paste("Reached convergence ratio of ", ansM3$tconv.max, sep = ""))
    break
  }
}

gofbeh <- sienaGOF(ansM3,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "freq_run")

gofbeh2 <- sienaGOF(ansM3,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "time_run")

goflist3 <- list(gofbeh, gofbeh2)
save(goflist3, file= paste("files", "/", "other clubs", "/", "club 2", "/", "gof3.RData", sep=""))

# save the sienaFit object
save(ansM3, file= paste("files", "/", "other clubs", "/", "club 2", "/", "model3.RData", sep=""))
```

<br>

**Violin plot of running frequency**
```{r class.source = 'fold-hide'}
load("files/other clubs/club 2/gof3.RData")
plot(goflist3[[1]])
```

**Violin plot of running duration**
```{r class.source = 'fold-hide'}
plot(goflist3[[2]])
```

#### Results {.tabset .tabset-fade}

##### Model 1
```{r, echo=F}
htmltools::includeHTML("files/other clubs/club 2/Model 2.html")
```

<!--- 

@RF: 
behavior dynamics: covariate effects (other sports and cross-effect) only in duration dynamics, not in frequency.
Also: all convergence t ratios <.11, not .10...
---> 
 

##### Model 2
```{r, echo=F}
htmltools::includeHTML("files/other clubs/club 2/Model 3.html")
```

<!--- 

@RF: 
Here it is better. I altered the code for myeff3 and myeff4.

---> 
 

#### {-}

### Club 3

First, we model network structure and dynamics in kudos tie formation. 

Club 3 is very large (N=165). We use only the first 3 waves to speed up the model optimization. When we find a satisfactory solution, we run the models on all waves.

First, we create a new R-Siena object for this club.

```{r echo=T, results='hide'}
load("clubdata.RData") # list with club data
club <- clubdata[[3]]

# create RSIENA object
kudonet <- sienaDependent(club$kudo[,,1:3], allowOnly = FALSE)
time_run <- sienaDependent(club$time_run[,,1:3], type= "behavior", allowOnly = FALSE)
freq_run <- sienaDependent(club$freq_run[,,1:3], type= "behavior", allowOnly = FALSE)
  
time_other <- varCovar(club$time_other[,,1:3])
freq_other <- varCovar(club$freq_other[,,1:3])

gender <- NA 
gender <- ifelse(club$male == 1, 1, gender)
gender <- ifelse(club$female == 1, 2, gender)
gender <- ifelse(club$other == 1, 2, gender)
gender <- coCovar(gender)

mydata <- sienaDataCreate(kudonet, time_run, freq_run, time_other, freq_other, gender)
```

<br>

First, we model network structure and dynamics in kudos tie formation. 

We include the following structural effects: *outdegree*, *reciprocity*, *transitivity*. Also, we added degree-related endogenous effects: *outdegree-activity* and *indegree-popularity*.

In this club, the *outIsolate*-effect (leading to now giving kudos to anyone) was rather important: a score-type test indicated that it was not 0 (not shown).

This rather simple model specification did not produce sufficient GOF (not shown).
GOF based on indegree- (p=0.031) and outdegree-distribution (p=0.007) was poor; GOF based on geodesic distance-distribution was all right.

We investigate how the model may be extended to obtain a better fit. We do this by a theory-guided exploration. We include the following effects: 

- effects that model the inverse indication of local hierarchy (*3-cycle* and *reciprocity x GWESP*)
- the *reciprocal degree-related activity* effect (reciAct; for which we expect a negative parameter estimate)
- an effect to model the relationsip between in- and outdegrees (*outdegree-popularity*)

Based on this, we do a foward model-selection and use the score-type test to see which of these effects could improve the model (spoiler: all of them do!). 

We use *sienaGOF* to estimate how the fit (based on Mahalanobis distance) would increase had we included the fixed-and-tested effects (hence, circumventing the need for re-estimating the model with the new effect).

```{r echo=T, results='hide'}
myeff <- getEffects(mydata) # define myeff object

# structural network effects
myeff1 <- includeEffects(myeff, gwespFF, name = "kudonet") 
myeff1 <- includeEffects(myeff1, outActSqrt, inPopSqrt, name = "kudonet") 
myeff1 <- includeEffects( myeff1, outIso, name = "kudonet")

# we fix-and-test these effects:

# 1. interaction GWESP x recip.
myeff1 <- includeInteraction(myeff1, recip, gwespFF, parameter = 69, name = "kudonet")
myeff1 # the interaction-effect is 9
eff1 <- myeff1[myeff1$include,]$effect1[9]
eff2 <- myeff1[myeff1$include,]$effect2[9]
myeff1 <- setEffect(myeff1, unspInt, fix=TRUE, test=TRUE,
                                    effect1=eff1, effect2=eff2)
# 2. 3-cycles effect
myeff1 <- setEffect(myeff1, cycle3, name = "kudonet", fix=TRUE, test=TRUE, include=TRUE)
# 3. reciprocal degree-related activity effect
myeff1 <- setEffect(myeff1, reciAct, name = "kudonet", fix=TRUE, test=TRUE, include=TRUE)
# 4. outdegree popularity effect
myeff1 <- setEffect( myeff1, outPopSqrt, name = "kudonet", fix=TRUE, test=TRUE, include=TRUE)

# selection effects for kudos ties formation based on running activity
myeff2 <- setEffect(myeff1, higher, name = "kudonet", interaction1 = "freq_run")

# and selection based on gender 
myeff2 <- includeEffects(myeff2, egoX, altX, sameX, name="kudonet", interaction1 = "gender" )
```

<br>

We (1) define the algorithm, (2) estimate the model, and (3) assess GOF based on three auxiliary network statistics: outdegree, indegree and geodesic distance distribution. We investigate how GOF based on these diagnostics would improve had we included the fixed-and-tested effects. Note that these effects are not controlled for each other, only for the estimated parameters. 

```{r eval=F, results='hide'}
myalgorithm <- sienaAlgorithmCreate(projname = "test") # set the SAOM algorithm 

ansM1 <- siena07(myalgorithm, data = mydata, effects = myeff2, # estimate the SAOM; 
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE) # if the previous (simple) model was also estimated, you can use 'prevAns=TRUE'

# we can check the score-test over the 4-degree of freedom aggregate.
# and seperately, by calling:
summary(ansM1)

# GOF diagnostics:
gofi <- sienaGOF(ansM1, 
                 IndegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

plot(gofi)

gofo <- sienaGOF(ansM1, 
                 OutdegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

gofgeo <- sienaGOF(ansM1, 
                 GeodesicDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

goflist <- list(gofi, gofo, gofgeo)
save(goflist, file= paste("files", "/", "other clubs", "/", "club 3", "/", "gof1.RData", sep=""))
```

#### {.tabset .tabset-fade}

##### Indegree distribution
```{r class.source = 'fold-hide'}
load("files/other clubs/club 3/gof1.RData")
plot(goflist[[1]])
summary(goflist[[1]])
```

<br>

MHD is the predicted Mahalanobis distance in the extended model (including the particular effect). Indegree GOF is predicted to increase the most if we include the 3-cycles effect.

<br>

----

##### Outdegree distribution
```{r class.source = 'fold-hide'}
plot(goflist[[2]])
summary(goflist[[2]])
```

<br>

We see that the Outdegree-distribution would improve if we include the interaction between reciprocity and transitivity (GWESP).

<br>

----

##### Geodesic distance distribution
```{r class.source = 'fold-hide'}
plot(goflist[[3]])
summary(goflist[[3]])
```

<br>

The recip x GWESP-effect seems to be the most promising: it is predicted to improve both in- and outdegree GOF (and also geodesic distances!).

<br>

----


#### {-}


We re-estimate the model with with the additional reciprocity-GWESP interaction-effect. We use the prevAns-option to use previous model-results as starting values for the estimation.

```{r eval=F, results='hide'}
myeff2 <- includeInteraction(myeff2, recip, gwespFF, parameter = 69, name = "kudonet")

ansM1 <- siena07(myalgorithm, data = mydata, effects = myeff2, prevAns = ansM1, # prior estimates as starting values
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE)

# GOF diagnostics:
gofi <- sienaGOF(ansM1, 
                 IndegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

gofo <- sienaGOF(ansM1, 
                 OutdegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

gofgeo <- sienaGOF(ansM1, 
                 GeodesicDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

goflist <- list(gofi, gofo, gofgeo)
save(goflist, file= paste("files", "/", "other clubs", "/", "club 3", "/", "gof1.2.RData", sep=""))
```

#### {.tabset .tabset-fade}

##### Indegree distribution
```{r class.source = 'fold-hide'}
load("files/other clubs/club 3/gof1.2.RData")
plot(goflist[[1]])
summary(goflist[[1]])
```

<br> 

Indegree GOF actually decreased. Modelling the relationship between in- and outdegree - through the *outPopSqrt* - may be a solution.

<br>

----

##### Outdegree distribution
```{r class.source = 'fold-hide'}
plot(goflist[[2]])
```

<br>

The fit for the outdegree distribution is now good!

<br>

----

##### Geodesic distance distribution
```{r class.source = 'fold-hide'}
plot(goflist[[3]])
```

<br>

Geodesic distance distribution GOF increased!

<br>

----


#### {-}

Let's  include the outPopSqrt-effect, to see how this improves indegree distribution GOF. We re-estimate the model with the new effect.

```{r eval=F, results='hide'}
myeff2 <- includeEffects(myeff2, outPopSqrt, name = "kudonet")

ansM1 <- siena07(myalgorithm, data = mydata, effects = myeff2, prevAns = ansM1, # prior estimates as starting values
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE)

# GOF diagnostics:
gofi <- sienaGOF(ansM1, 
                 IndegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

gofo <- sienaGOF(ansM1, 
                 OutdegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

gofgeo <- sienaGOF(ansM1, 
                 GeodesicDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

goflist <- list(gofi, gofo, gofgeo)
save(goflist, file= paste("files", "/", "other clubs", "/", "club 3", "/", "gof1.3.RData", sep=""))
```

#### {.tabset .tabset-fade}

##### Indegree distribution
```{r class.source = 'fold-hide'}
load("files/other clubs/club 3/gof1.3.RData")
plot(goflist[[1]])
```


The fit for the indegree distribution is now good!

<br>

----

##### Outdegree distribution
```{r class.source = 'fold-hide'}
plot(goflist[[2]])
```

Outdegree GOF is still all right.

<br>

----

##### Geodesic distance distribution
```{r class.source = 'fold-hide'}
plot(goflist[[3]])
```

GD GOF is still all right.


<br>

----

#### {-}


Now that we reached adequate GOF for the kudos-network, it is time to model running activity dynamics:

- interdepence between frequency and duration of running.
- interdependence between running and other sports.
- gender-effects on behavior change.
- two influence effects: *indegree* and *upward assimilation*. We control for the outdegree-effect: we fix it and score-type test it.

Again, we face convergence issues due to overfitting. We fix effects on behavior change of gender and other sports to 0 (score-type tests tell that these are 0).

```{r echo=T, results='hide'}
myeff3 <- includeEffects(myeff2, effFrom, name = "time_run", interaction1 = "freq_run")
myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "time_run")
myeff3 <- includeEffects(myeff3, effFrom, name = "time_run", interaction1 = "time_other")
myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "freq_other")

# gender effects fixed-and-tested
myeff3 <- setEffect(myeff3, effFrom, name = "freq_run", interaction1 = "gender", fix=T, test = F, initialValue = 0)
myeff3 <- setEffect(myeff3, effFrom, name = "time_run", interaction1 = "gender", fix=T, test = F, initialValue = 0)
# same for 'other sports'
myeff3 <- setEffect(myeff3, effFrom, name = "time_run", interaction1 = "time_other", fix=T, test=F, initialValue=0)
myeff3 <- setEffect(myeff3, effFrom, name = "freq_run", interaction1 = "freq_other", fix=T, test=F, initialValue=0)

# influence effects:
# 1. indegree effect
myeff3 <- includeEffects(myeff3, indeg, name = "freq_run", interaction1 = "kudonet") 
myeff3 <- includeEffects(myeff3, indeg, name = "time_run", interaction1 = "kudonet")
# 2. upward assimilation
myeff4 <- includeEffects(myeff3, avAttHigher, name = "freq_run", interaction1 = "kudonet") 
myeff4 <- includeEffects(myeff4, avAttHigher, name = "time_run", interaction1 = "kudonet")

# outdegree effect (fixed-and-tested)
myeff4 <- setEffect(myeff4, outdeg, name = "freq_run", interaction1 = "kudonet", fix=T, test=F, initialValue=0)
myeff4 <- setEffect(myeff4, outdeg, name = "time_run", interaction1 = "kudonet", fix=T, test=F, initialValue=0)
```


<br> 

We estimate the model and investigate another GOF-diagnostic: the distribution of running behavior variables. We estimate (1) a model with the objective function for running activity comprising the basic shape effects, covariate effects, and *indegree* effect; and (2) a model including the *upward assimilation* and *outdegree* effect. 

We run these models until we reach convergence. 

#### {.tabset .tabset-fade}

##### The indegree-effect
```{r eval=F, results='hide'}
try <- 1
ansM2 <- siena07(myalgorithm, data = mydata, effects = myeff3, # estimate the SAOM
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE, prevAns=ansM1)

# the following script lets the model re-run until we get a good convergence ratio
while (TRUE){
  
  if(ansM2$tconv.max >= .25){
    try <- try + 1
    print(paste("Model did not converge: ", ansM2$tconv.max, sep = ""))
    print(paste("Try:", try, sep=" "))
    ansM2 <- siena07( myalgorithm, data = mydata, effects = myeff3, prevAns= ansM2, returnDeps=TRUE)
    
  }else{
    siena.table(ansM2, type="html", tstat=T, d=3, sig=T, file = paste("files", "/", "other clubs", "/", "club 3", "/", "Model 2", ".html", sep=""))
    print(paste("Reached convergence ratio of ", ansM2$tconv.max, sep = ""))
    break
  }
}

gofbeh <- sienaGOF(ansM2,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "freq_run")

gofbeh2 <- sienaGOF(ansM2,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "time_run")

goflist2 <- list(gofbeh, gofbeh2)
save(goflist2, file= paste("files", "/", "other clubs", "/", "club 3", "/", "gof2.RData", sep=""))

# save the sienaFit object
save(ansM2, file= paste("files", "/", "other clubs", "/", "club 3", "/", "model2.RData", sep=""))
```

<br>

**Violin plot of running frequency**
```{r class.source = 'fold-hide'}
load("files/other clubs/club 3/gof2.RData")
plot(goflist2[[1]])
```

**Violin plot of running duration**
```{r class.source = 'fold-hide'}
plot(goflist2[[2]])
```

Acceptable GOF for frequency, not for duration...

##### The upward assimilation effect


```{r eval=F, results='hide'}
try <- 1
ansM3 <- siena07(myalgorithm, data = mydata, effects = myeff4, # estimate the SAOM
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE, prevAns=ansM2)

# the following script lets the model re-run until we get a good convergence ratio
while (TRUE){
  
  if(ansM3$tconv.max >= .25){
    try <- try + 1
    print(paste("Model did not converge: ", ansM3$tconv.max, sep = ""))
    print(paste("Try:", try, sep=" "))
    ansM3 <- siena07( myalgorithm, data = mydata, effects = myeff4, prevAns= ansM3, returnDeps=TRUE)
    
  }else{
    siena.table(ansM3, type="html", tstat=T, d=3, sig=T, file = paste("files", "/", "other clubs", "/", "club 3", "/", "Model 3", ".html", sep=""))
    print(paste("Reached convergence ratio of ", ansM3$tconv.max, sep = ""))
    break
  }
}

gofbeh <- sienaGOF(ansM3,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "freq_run")

gofbeh2 <- sienaGOF(ansM3,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "time_run")

goflist3 <- list(gofbeh, gofbeh2)
save(goflist3, file= paste("files", "/", "other clubs", "/", "club 3", "/", "gof3.RData", sep=""))

# save the sienaFit object
save(ansM3, file= paste("files", "/", "other clubs", "/", "club 3", "/", "model3.RData", sep=""))
```

<br>

**Violin plot of running frequency**
```{r class.source = 'fold-hide'}
load("files/other clubs/club 4/gof3.RData")
plot(goflist3[[1]])
```

**Violin plot of running duration**
```{r class.source = 'fold-hide'}
plot(goflist3[[2]])
```

#### {-}

<br>

----

This model configuration seems promising. Let's estimate the model on all waves. This will take a long time! We re-estimate the models until we reach convergence. We save the sienaFit-objects of both models for subsequent meta-analyses.


```{r eval=F, results='hide', class.source = 'fold-hide'}
mydata <- clubdata_rsiena[[3]] # grab club
myeff <- getEffects(mydata) # define myeff object

myeff1 <- includeEffects(myeff, gwespFF, outActSqrt, inPopSqrt, outPopSqrt, outIso, name = "kudonet") 
myeff1 <- includeInteraction(myeff1, recip, gwespFF, parameter = 69, name = "kudonet")
myeff1 <- setEffect(myeff1, reciAct, name = "kudonet", fix=TRUE, test=FALSE, initialValue = 0)

myeff2 <- setEffect(myeff1, higher, name = "kudonet", interaction1 = "freq_run")
myeff2 <- includeEffects(myeff2, egoX, altX, sameX, name="kudonet", interaction1 = "gender" )

myeff3 <- includeEffects(myeff2, effFrom, name = "time_run", interaction1 = "freq_run")
myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "time_run")
myeff3 <- setEffect(myeff3, effFrom, name = "time_run", interaction1 = "time_other", fix=TRUE, test=FALSE, initialValue = 0)
myeff3 <- setEffect(myeff3, effFrom, name = "freq_run", interaction1 = "freq_other", fix=TRUE, test=FALSE, initialValue = 0)
myeff3 <- setEffect(myeff3, effFrom, name = "freq_run", interaction1 = "gender", fix=TRUE, test = FALSE, initialValue = 0)
myeff3 <- setEffect(myeff3, effFrom, name = "time_run", interaction1 = "gender", fix=TRUE, test = FALSE, initialValue = 0)

myeff3 <- includeEffects(myeff3, indeg, name = "freq_run", interaction1 = "kudonet") 
myeff3 <- includeEffects(myeff3, indeg, name = "time_run", interaction1 = "kudonet")

myeff4 <- includeEffects(myeff3, avAttHigher, name = "freq_run", interaction1 = "kudonet") 
myeff4 <- includeEffects(myeff4, avAttHigher, name = "time_run", interaction1 = "kudonet")

myeff4 <- setEffect(myeff4, outdeg, name = "freq_run", interaction1 = "kudonet", fix=TRUE, test=FALSE, initialValue=0)
myeff4 <- setEffect(myeff4, outdeg, name = "time_run", interaction1 = "kudonet", fix=TRUE, test=FALSE, initialValue=0)

# set the SAOM algorithm 
myalgorithm <- sienaAlgorithmCreate(projname = "test") 
## Estimate model with indegree effect
try <- 1

ansM2 <- siena07(myalgorithm, data = mydata, effects = myeff3, # estimate the SAOM
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE)

# the following script lets the model re-run until we get a good convergence ratio
while (TRUE){
  
  if(ansM2$tconv.max >= .25){
    try <- try + 1
    print(paste("Model did not converge: ", ansM2$tconv.max, sep = ""))
    print(paste("Try:", try, sep=" "))
    ansM2 <- siena07( myalgorithm, data = mydata, effects = myeff3, prevAns= ansM2, returnDeps=TRUE)
    
  }else{
    siena.table(ansM2, type="html", tstat=T, d=3, sig=T, file = paste("files", "/", "other clubs", "/", "club 3", "/", "Model 2 full", ".html", sep=""))
    print(paste("Reached convergence ratio of ", ansM2$tconv.max, sep = ""))
    break
  }
}

# GOF assessment
gofbeh <- sienaGOF(ansM2,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "freq_run")

gofbeh2 <- sienaGOF(ansM2,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "time_run")

goflist2 <- list(gofbeh, gofbeh2)
save(goflist2, file= paste("files", "/", "other clubs", "/", "club 3", "/", "gof2_full.RData", sep=""))

# save the sienaFit object
save(ansM2, file= paste("files", "/", "other clubs", "/", "club 3", "/", "model2_full.RData", sep=""))

## Estimate the next model including the influence effect:
try <- 1

ansM3 <- siena07(myalgorithm, data = mydata, effects = myeff4, # estimate the SAOM
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE)

# the following script lets the model re-run until we get a good convergence ratio
while (TRUE){
  
  if(ansM3$tconv.max >= .25){
    try <- try + 1
    print(paste("Model did not converge: ", ansM3$tconv.max, sep = ""))
    print(paste("Try:", try, sep=" "))
    ansM3 <- siena07( myalgorithm, data = mydata, effects = myeff4, prevAns= ansM3, returnDeps=TRUE)
    
  }else{
    siena.table(ansM3, type="html", tstat=T, d=3, sig=T, file = paste("files", "/", "other clubs", "/", "club 3", "/", "Model 3 full", ".html", sep=""))
    print(paste("Reached convergence ratio of ", ansM3$tconv.max, sep = ""))
    break
  }
}

# GOF assessment
gofbeh <- sienaGOF(ansM3,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "freq_run")

gofbeh2 <- sienaGOF(ansM3,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "time_run")

goflist2 <- list(gofbeh, gofbeh2)
save(goflist2, file= paste("files", "/", "other clubs", "/", "club 3", "/", "gof3_full.RData", sep=""))

# save the sienaFit object
save(ansM2, file= paste("files", "/", "other clubs", "/", "club 3", "/", "model3_full.RData", sep=""))
```

<br>

#### Results {.tabset .tabset-fade}

##### Model 1
```{r, echo=F}
htmltools::includeHTML("files/other clubs/club 3/Model 2.html")
```

##### Model 2
```{r, echo=F}
htmltools::includeHTML("files/other clubs/club 3/Model 3.html")
```

### Club 4

First, we model network structure and dynamics in kudos tie formation. 

In this club, the *outIsolate*-effect (leading to now giving kudos to anyone) was rather important: a score-type test indicated that it was not 0 (not shown).
Selection effects of gender and running (monadic and dyadic) were also score-type tested and fixed to 0.

```{r echo=T, results='hide'}
mydata <- clubdata_rsiena[[4]] # grab club

myeff <- getEffects(mydata) # define myeff object

# structural network effects
myeff1 <- includeEffects(myeff, gwespFF, name = "kudonet") 
myeff1 <- includeEffects(myeff1, outActSqrt, inPopSqrt, name = "kudonet") 
myeff1 <- setEffect( myeff1, outIso, name = "kudonet")

# selection effect for kudos ties formation based on running activity
myeff2 <- setEffect( myeff1, higher, name = "kudonet", interaction1 = "freq_run")

# and selection based on gender (fixed to 0)
myeff2 <- setEffect( myeff2, egoX, name = "kudonet", interaction1 = "gender", fix=TRUE, test=FALSE, initialValue = 0)
myeff2 <- setEffect( myeff2, altX, name = "kudonet", interaction1 = "gender", fix=TRUE, test=FALSE, initialValue = 0)
myeff2 <- setEffect( myeff2, sameX, name = "kudonet", interaction1 = "gender", fix=TRUE, test=FALSE, initialValue = 0)
```

<!--- 

@RF: To include the selection on behavior effects:
myeff2 <- includeEffects(myeff1, egoX, altX, simX, name = "kudonet", interaction1 = "freq_run")
To include gender-effects:
myeff2 <- includeEffects(myeff2, egoX, altX, sameX, name="kudonet", interaction1 = "gender" )

---> 

<br>

We (1) define the algorithm, (2) estimate the model, and (3) assess GOF based on three auxiliary network statistics: otudegree, indegree and geodesic distance distribution. We will run the model as many times as necessary, until we reach a convergence ratio of < .25.

```{r eval=F, results='hide'}
myalgorithm <- sienaAlgorithmCreate(projname = "test") # set the SAOM algorithm 

try <- 1
ansM1 <- siena07(myalgorithm, data = mydata, effects = myeff2, # estimate the SAOM
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE)

# the following script lets the model re-run until we get a good convergence ratio
while (TRUE){
  
  if(ansM1$tconv.max >= .25){
    try <- try + 1
    print(paste("Model did not converge: ", ansM1$tconv.max, sep = ""))
    print(paste("Try:", try, sep=" "))
    ansM1 <- siena07( myalgorithm, data = mydata, effects = myeff2, prevAns= ansM1, returnDeps=TRUE)
    
  }else{
    print(paste("Reached overall maximum convergence ratio of ", ansM1$tconv.max, sep = ""))
    break
  }
}

gofi <- sienaGOF(ansM1, 
                 IndegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

gofo <- sienaGOF(ansM1, 
                 OutdegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

GeodesicDistribution <- function (i, data, sims, period, groupName,
   varName, levls=c(1:5, Inf), cumulative=TRUE, ...) {
     x <- networkExtraction(i, data, sims, period, groupName, varName)
     require(sna)
     a <- sna::geodist(symmetrize(x))$gdist
     if (cumulative)
     {
       gdi <- sapply(levls, function(i){ sum(a<=i) })
     }
     else
     {
       gdi <- sapply(levls, function(i){ sum(a==i) })
     }
     names(gdi) <- as.character(levls)
     gdi
}

gofgeo <- sienaGOF(ansM1, 
                 GeodesicDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

goflist <- list(gofi, gofo, gofgeo)
save(goflist, file= paste("files", "/", "other clubs", "/", "club 4", "/", "gof1.RData", sep=""))
```

#### {.tabset .tabset-fade}

##### Indegree distribution
```{r class.source = 'fold-hide'}
load("files/other clubs/club 4/gof1.RData")
plot(goflist[[1]])
```

##### Outdegree distribution
```{r class.source = 'fold-hide'}
plot(goflist[[2]])
```

##### Geodesic distance distribution
```{r class.source = 'fold-hide'}
plot(goflist[[3]])
```

#### {-}


GOF is all right!

<br>

We fine-tune the network-part of the model: we include some parameters as fixed to 0 for subsequent meta-analysis:

```{r echo=T, results='hide'}
myeff2 <- setEffect(myeff2, outPopSqrt, name = "kudonet", fix=TRUE, test=FALSE, initialValue = 0)
myeff2 <- setEffect(myeff2, reciAct, name = "kudonet", fix=TRUE, test=FALSE, initialValue = 0)
(myeff2 <- includeInteraction(myeff2, recip, gwespFF, parameter = 69, name = "kudonet"))
(eff1 <- myeff2[myeff2$include,]$effect1[24])
(eff2 <- myeff2[myeff2$include,]$effect2[24])
myeff2 <- setEffect(myeff2, unspInt, fix=TRUE, test=FALSE,
                                    effect1=eff1, effect2=eff2)
```


<br> 

Second, we model dynamics in running behaviors. 

```{r echo=T, results='hide'}
# covariate effects:
# interdependency frequency and duration
myeff3 <- includeEffects(myeff2, effFrom, name = "time_run", interaction1 = "freq_run")
myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "time_run")
# effects from other sports
myeff3 <- includeEffects(myeff3, effFrom, name = "time_run", interaction1 = "time_other")
myeff3 <- includeEffects(myeff3, effFrom, name = "freq_run", interaction1 = "freq_other")
# gender effects (fixed to 0)
myeff3 <- setEffect(myeff3, effFrom, name = "freq_run", interaction1 = "gender", fix=T, test=F, initialValue=0)
myeff3 <- setEffect(myeff3, effFrom, name = "time_run", interaction1 = "gender", fix=T, test=F, initialValue=0)

# influence effects:

# 1. indegree effect
myeff3 <- includeEffects(myeff3, indeg, name = "freq_run", interaction1 = "kudonet") 
myeff3 <- includeEffects(myeff3, indeg, name = "time_run", interaction1 = "kudonet")
# 2. upward assimilation
myeff4 <- includeEffects(myeff3, avAttHigher, name = "freq_run", interaction1 = "kudonet") 
myeff4 <- includeEffects(myeff4, avAttHigher, name = "time_run", interaction1 = "kudonet")

# 3. downward assimilation
#myeff4 <- includeEffects(myeff4, avAttLower, name = "freq_run", interaction1 = "kudonet") 
#myeff4 <- includeEffects(myeff4, avAttLower, name = "time_run", interaction1 = "kudonet")

# outdegree effect (tested and fixed at 0)
myeff4 <- setEffect(myeff4, outdeg, name = "freq_run", interaction1 = "kudonet", fix=T, test=F, initialValue=0)
myeff4 <- setEffect(myeff4, outdeg, name = "time_run", interaction1 = "kudonet", fix=T, test=F, initialValue=0)

# rate parameter at fixed value
myeff4 <- setEffect(myeff4, Rate, name = "freq_run", fix = TRUE, test = FALSE, type = "rate", period = 8, initialValue = 5  )
```


<br> 

We estimate the model and investigate another GOF-diagnostic: the distribution of running behavior variables. We estimate (1) a model with the objective function for running activity comprising the basic shape effects, covariate effects, and *indegree* effect; and (2) a model including the *upward assimilation* and *outdegree* effect. We save the sienaFit-objects of both models for subsequent meta-analyses.

This club faced convergence issues in the last model. We achieved convergence by (1) fixing the outdegree-effect to 0 (score-type test indicated that this effect was 0) and fixing the rate parameter for running frequency dynamics to 5 in period 8 (we aimed at a *t*-ratio for convergence for this parameter of <.2).

#### {.tabset .tabset-fade}

##### The indegree-effect
```{r eval=F, results='hide'}
try <- 1
ansM2 <- siena07(myalgorithm, data = mydata, effects = myeff3, # estimate the SAOM
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE)

# the following script lets the model re-run until we get a good convergence ratio
while (TRUE){
  
  if(ansM2$tconv.max >= .25){
    try <- try + 1
    print(paste("Model did not converge: ", ansM2$tconv.max, sep = ""))
    print(paste("Try:", try, sep=" "))
    ansM2 <- siena07( myalgorithm, data = mydata, effects = myeff3, prevAns= ansM2, returnDeps=TRUE)
    
  }else{
    siena.table(ansM2, type="html", tstat=T, d=3, sig=T, file = paste("files", "/", "other clubs", "/", "club 4", "/", "Model 2", ".html", sep=""))
    print(paste("Reached overall maximum convergence ratio of ", ansM2$tconv.max, sep = ""))
    break
  }
}

gofbeh <- sienaGOF(ansM2,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "freq_run")

gofbeh2 <- sienaGOF(ansM2,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "time_run")

goflist2 <- list(gofbeh, gofbeh2)
save(goflist2, file= paste("files", "/", "other clubs", "/", "club 4", "/", "gof2.RData", sep=""))

# save the sienaFit object
save(ansM2, file= paste("files", "/", "other clubs", "/", "club 4", "/", "model2.RData", sep=""))
```

<br>

**Violin plot of running frequency**
```{r class.source = 'fold-hide'}
load("files/other clubs/club 4/gof2.RData")
plot(goflist2[[1]])
```

**Violin plot of running duration**
```{r class.source = 'fold-hide'}
plot(goflist2[[2]])
```


##### The upward assimilation effect
```{r eval=F, results='hide'}
try <- 1
ansM3 <- siena07(myalgorithm, data = mydata, effects = myeff4, # estimate the SAOM
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE)

# the following script lets the model re-run until we get a good convergence ratio
while (TRUE){
  
  if(ansM3$tconv.max >= .25){
    try <- try + 1
    print(paste("Model did not converge: ", ansM3$tconv.max, sep = ""))
    print(paste("Try:", try, sep=" "))
    ansM3 <- siena07( myalgorithm, data = mydata, effects = myeff4, prevAns= ansM3, returnDeps=TRUE)
    
  }else{
    siena.table(ansM3, type="html", tstat=T, d=3, sig=T, file = paste("files", "/", "other clubs", "/", "club 4", "/", "Model 3", ".html", sep=""))
    print(paste("Reached convergence ratio of ", ansM3$tconv.max, sep = ""))
    break
  }
}

gofbeh <- sienaGOF(ansM3,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "freq_run")

gofbeh2 <- sienaGOF(ansM3,
                   BehaviorDistribution,
                   verbose = TRUE,
                   join = TRUE,
                   varName = "time_run")

goflist3 <- list(gofbeh, gofbeh2)
save(goflist3, file= paste("files", "/", "other clubs", "/", "club 4", "/", "gof3.RData", sep=""))

# save the sienaFit object
save(ansM3, file= paste("files", "/", "other clubs", "/", "club 4", "/", "model3.RData", sep=""))
```

<br>

**Violin plot of running frequency**
```{r class.source = 'fold-hide'}
load("files/other clubs/club 4/gof3.RData")
plot(goflist3[[1]])
```

**Violin plot of running duration**
```{r class.source = 'fold-hide'}
plot(goflist3[[2]])
```

#### {-}

#### Results {.tabset .tabset-fade}

##### Model 1
```{r, echo=F}
htmltools::includeHTML("files/other clubs/club 4/Model 2.html")
```

##### Model 2
```{r, echo=F}
htmltools::includeHTML("files/other clubs/club 4/Model 3.html")
```

#### {-}

### Club 5

First, we model network structure and dynamics in kudos tie formation. 

We include the following structural effects: *outdegree*, *reciprocity*, *transitivity*. Also, we added degree-related endogenous effects: *outdegree-activity* and *indegree-popularity*.

In this club, the *outIsolate*-effect (leading to now giving kudos to anyone) was rather important: a score-type test indicated that it was not 0 (not shown).

This rather simple model specification did not produce sufficient GOF: p=0 for all 3 GOF diagnostics. 

We investigate how the model may be extended to obtain a better fit. We do this by a theory-guided exploration. We include the following effects: 

- effects that model the inverse indication of local hierarchy (*3-cycle* and *reciprocity x GWESP*)
- the *reciprocal degree-related activity* effect (reciAct; for which we expect a negative parameter estimate)
- an effect to model the relationsip between in- and outdegrees (*outdegree-popularity*)

Based on this, we do a foward model-selection and use the score-type test to see which of these effects could improve the model (spoiler: all of them do!). 

We use *sienaGOF* to estimate how the fit (based on Mahalanobis distance) would increase had we included the fixed-and-tested effects (hence, circumventing the need for re-estimating the model with the new effect). 

<!--- 

@RF: 
in all networks, out- and indegrees are strongly dispersed. So we may also want to let the rate function condition on the log-outdegree (outRateLog)

---> 
 

```{r echo=T, results='hide'}
mydata <- clubdata_rsiena[[5]] # grab club

myeff <- getEffects(mydata) # define myeff object

# structural network effects
myeff1 <- includeEffects(myeff, gwespFF, name = "kudonet") 
myeff1 <- includeEffects(myeff1, outActSqrt, inPopSqrt, name = "kudonet") 
myeff1 <- includeEffects( myeff1, outIso, name = "kudonet")

# we fix-and-test these effects:

# 1. interaction GWESP x recip.
myeff1 <- includeInteraction(myeff1, recip, gwespFF, parameter = 69, name = "kudonet")
myeff1 # the interaction-effect is 18
eff1 <- myeff1[myeff1$include,]$effect1[18]
eff2 <- myeff1[myeff1$include,]$effect2[18]
myeff1 <- setEffect(myeff1, unspInt, fix=TRUE, test=TRUE,
                                    effect1=eff1, effect2=eff2)
# 2. 3-cycles effect
myeff1 <- setEffect(myeff1, cycle3, name = "kudonet", fix=TRUE, test=TRUE, include=TRUE)
# 3. reciprocal degree-related activity effect
myeff1 <- setEffect(myeff1, reciAct, name = "kudonet", fix=TRUE, test=TRUE, include=TRUE)
# 4. outdegree popularity effect
myeff1 <- setEffect( myeff1, outPopSqrt, name = "kudonet", fix=TRUE, test=TRUE, include=TRUE)

# selection effects for kudos ties formation based on running activity
myeff2 <- setEffect(myeff1, higher, name = "kudonet", interaction1 = "freq_run")

# and selection based on gender 
myeff2 <- includeEffects(myeff2, egoX, altX, sameX, name="kudonet", interaction1 = "gender" )
```

<br>

We (1) define the algorithm, (2) estimate the model, and (3) assess GOF based on three auxiliary network statistics: outdegree, indegree and geodesic distance distribution. We investigate how GOF based on these diagnostics would improve had we included the fixed-and-tested effects. Note that these effects are not controlled for each other, only for the estimated parameters. 

```{r eval=F, results='hide'}
myalgorithm <- sienaAlgorithmCreate(projname = "test") # set the SAOM algorithm 

ansM1 <- siena07(myalgorithm, data = mydata, effects = myeff1, # estimate the SAOM
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE)

# we can check the score-test over the 4-degree of freedom aggregate.
# and seperately, by calling:
summary(ansM1)

# GOF diagnostics:
gofi <- sienaGOF(ansM1, 
                 IndegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

gofo <- sienaGOF(ansM1, 
                 OutdegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

GeodesicDistribution <- function (i, data, sims, period, groupName,
   varName, levls=c(1:5, Inf), cumulative=TRUE, ...) {
     x <- networkExtraction(i, data, sims, period, groupName, varName)
     require(sna)
     a <- sna::geodist(symmetrize(x))$gdist
     if (cumulative)
     {
       gdi <- sapply(levls, function(i){ sum(a<=i) })
     }
     else
     {
       gdi <- sapply(levls, function(i){ sum(a==i) })
     }
     names(gdi) <- as.character(levls)
     gdi
}

gofgeo <- sienaGOF(ansM1, 
                 GeodesicDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

goflist <- list(gofi, gofo, gofgeo)
save(goflist, file= paste("files", "/", "other clubs", "/", "club 5", "/", "gof1.RData", sep=""))
```

#### {.tabset .tabset-fade}

##### Indegree distribution
```{r class.source = 'fold-hide'}
load("files/other clubs/club 5/gof1.RData")
plot(goflist[[1]])
summary(goflist[[1]])
```

<br>

Indegree GOF is bad (p=0). MHD is the predicted Mahalanobis distance in the extended model (including the particular effect). Our fixed-and-tested effects are predicted to only marginally improve the fit of the indegree-distribution. Most promising is reciAct, closely followed by outPopSqrt. 

<br>

----

##### Outdegree distribution
```{r class.source = 'fold-hide'}
plot(goflist[[2]])
summary(goflist[[2]])
```

<br>

Outdegree GOF is bad (p=0). MHD is the predicted Mahalanobis distance in the extended model (including the particular effect). We see that the Outdegree-distribution is predicted to improve drastically if we include the interaction between reciprocity and transitivity (GWESP).

<br>

----

##### Geodesic distance distribution
```{r class.source = 'fold-hide'}
plot(goflist[[3]])
summary(goflist[[3]])
```

<br>

GD-distribution fit would improve most after including the interaction between reciprocity and transitivity (GWESP). This effect seems most promising for improving GOF diagnostics.

<br>

----


#### {-}

We re-estimate the model with with the additional reciprocity-GWESP interaction-effect. We use the prevAns-option to use previous model-results as starting values for the estimation.

```{r eval=F, results='hide'}
myeff2 <- includeInteraction(myeff2, recip, gwespFF, parameter = 69, name = "kudonet")

ansM1 <- siena07(myalgorithm, data = mydata, effects = myeff2, prevAns = ansM1, # prior estimates as starting values
                 batch = FALSE, verbose = FALSE, returnDeps = TRUE)

# GOF diagnostics:
gofi <- sienaGOF(ansM1, 
                 IndegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

gofo <- sienaGOF(ansM1, 
                 OutdegreeDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

gofgeo <- sienaGOF(ansM1, 
                 GeodesicDistribution, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "kudonet")

goflist <- list(gofi, gofo, gofgeo)
save(goflist, file= paste("files", "/", "other clubs", "/", "club 5", "/", "gof1.2.RData", sep=""))
```

#### {.tabset .tabset-fade}

##### Indegree distribution
```{r class.source = 'fold-hide'}
load("files/other clubs/club 5/gof1.2.RData")
plot(goflist[[1]])
summary(goflist[[1]])
```

Indegree GOF is still bad (p=0).

<br>

----

##### Outdegree distribution
```{r class.source = 'fold-hide'}
plot(goflist[[2]])
summary(goflist[[2]])
```

Outdegree GOF is still bad (p=0).

<br>

----

##### Geodesic distance distribution
```{r class.source = 'fold-hide'}
plot(goflist[[3]])
summary(goflist[[3]])
```

<br>

GD GOF is still bad (p=0). 

<br>

----



## {-}

<br>

----

# Meta-analysis 

We summarize the results over the 5 clubs using Fisher's method for combining independent *p*-values. We perform a double test:

1. For detecting if there are any club-networks with a positive parameter value, the null hypothesis tested is: for all clubs, the value of this parameter is zero or less than zero; with the alternative hypothesis: for at least one club, the value of this parameter is greater than zero.
2. For detecting if there are any club-networks with a negative parameter value, the null hypothesis tested is: for all clubs, the value of this parameter is zero or greater than zero; with the alternative hypothesis: for at least one club, the value of this parameter is less than zero.

For each of these combined tests, the *p*-value is given. We use for each test the significance level of **/2, which yields an overall combined test at significance level **. Four different overall results are possible. Indicating the right-sided and left-sided *p*-values by *p<sub>r</sub>* and *p<sub>l</sub>* respectively, these results are:

- *p<sub>r</sub>* $\gt$ **/2 & *p<sub>l</sub>* $\gt$ **/2:

No evidence for any non-zero parameter values.

- *p<sub>r</sub>* $\leq$ **/2 & *p<sub>l</sub>* $\gt$ **/2:

Evidence that some club-networks have a positive parameter value, but no evidence for any negative parameter values.

- *p<sub>r</sub>* $\gt$ **/2 & *p<sub>l</sub>* $\leq$ **/2:

Evidence that some clubs have a negative parameter value, no evidence for any positive parameter values

- *p<sub>r</sub>* $\leq$ **/2 & *p<sub>l</sub>* $\leq$ **/2:

Evidence that some clubs have a negative parameter value, and some others have a positive parameter value.

<br>

----

First, we read in the sienaFit-objects of our 5 clubs. We use a function to load the objects under new names. Note: We saved 2 sienaFit-objects for each club.
Then we make a list of the sienaFit-objects of the 2 model specifications.

```{r eval=F}
# clean the working environment 
rm (list = ls( ))

loadRData <- function(fileName){
#loads an RData file, and returns it
    load(fileName)
    get(ls()[ls() != "fileName"])
}

club1_m1 <- loadRData("files/test club 1/model2.RData")
club2_m1 <- loadRData("files/other clubs/club 2/model2.RData")
club4_m1 <- loadRData("files/other clubs/club 4/model2.RData")

club1_m2 <- loadRData("files/test club 1/model3.RData")
club2_m2 <- loadRData("files/other clubs/club 2/model3.RData")
club4_m2 <- loadRData("files/other clubs/club 4/model3.RData")

m1List <- list(club1_m1, club2_m1, club4_m1)
m2List <- list(club1_m2, club2_m2, club4_m2)
```

<!--- 

@RF: just club 1 and 4 now, for testing purposes

---> 
<br>

We perform the meta-analysis seperately for the 2 model specifications. Our meta-analysis focuses only on the behavior dynamics.

<!--- 

@RF: for online supplementary material we will also present the meta-analysis for the structural and selection effects.

---> 

- Model 1: objective function for running activity comprising the basic shape effects, covariate effects, and indegree effect
- Model 2: including the upward assimilation effect and the outdegree effect

## {.tabset .tabset-fade}

### Model 1

We extract the parameter estimates and standard errors for the running activity objective functions.

```{r eval=F}
# {r results='hide'}
parameters <- sapply(m1List, function(x){x$theta[22:55]})
standers <- sapply(m1List, function(x){x$se[22:55]}) 

(eff.names <-
   m1List[[1]]$effects[m1List[[1]]$effects$include,'effectName']) 
(eff.names <- eff.names[22:55])

#combine parameters and std.errors with effect names
rownames(parameters) <- eff.names
rownames(standers) <- eff.names

# print rounded to 3 decimals
round(parameters,3)
round(standers,3)
```

<br>

We use *siena08* to perform Fisher's method for combining independent *p*-values.

<!--- 

@RF: 

- cjplus: test statistic for combination of right one-sided Fisher combination test
- cjplusp: p-value for cjplus
- cjminus: test statistic for combination of left one-sided Fisher combination test
- cjminusp: p-value for cjminus
- df: degrees of freedom

https://search.r-project.org/CRAN/refmans/RSiena/html/siena08.html 

Funnelplot?

---> 

```{r eval=F}
#{r results='hide'}
ans8 <- siena08(m1List, bound=100)


Overalls <- t(sapply(22:55, function(i){c(ans8[[i]]$Tsq, ans8[[i]]$pTsq)}))
round(Overalls, 3)

efnames <- names(ans8)[22:55]
efnames <- substring(efnames, 8)
rownames(Overalls) <- efnames
round(Overalls, 3)

Fishers <- t(sapply( 22:55,
                     function(i) {
                       c(ans8[[i]]$cjplus, ans8[[i]]$cjminus, ans8[[i]]$cjplusp, ans8[[i]]$cjminusp, 2 * ans8[[i]]$n1 )
                       }))
Fishers <- as.data.frame(Fishers)
rownames(Fishers) <- efnames
names(Fishers) <- c('Fplus', 'Fminus', 'pplus', 'pminus', 'df')
```

<br>

And we present the Fisher combinations:
```{r eval=F}
round(Fishers,3)
```

### Model 2

## {-}



<br>

----


### References 
